\author{Vladimir Bashkardin and Sergey Fomel}
\title{BP 2007 anisotropic velocity-analysis benchmark dataset}

\maketitle

\begin{abstract}
  This dataset was created by Hemang Shah and is provided at
  \url{http://software.seg.org/} courtesy of BP Exploration Operation
  Company Limited. It was released by Joe Dellinger and BP for the
  \href{http://www.eage.org/events/index.php?evp=2973&ActiveMenu=17&Opendivs=s3,s16,s17}{migration
    velocity-analysis} workshop at the 2010 EAGE meeting in
  Barcelona. The model is 2-D TTI (tilted transversely isotropic).
\end{abstract}

\section{model}
\inputdir{model}

\plot{vp}{width=\textwidth}{Vertical velocity.} \clearpage
\plot{vx}{width=\textwidth}{Horizontal velocity.} \clearpage
\plot{epsilon}{width=\textwidth}{Thomsen's $\epsilon$ parameter.} \clearpage
\plot{delta}{width=\textwidth}{Thomsen's $\delta$ parameter.} \clearpage
\plot{eta}{width=\textwidth}{Alkhalifah-Tsvankin's $\eta$ parameter.} \clearpage
\plot{theta}{width=\textwidth}{Tilt of the TI symmetry axis.} \clearpage

\plot{shots}{width=\textwidth}{Shots from finite-difference modeling.} \clearpage
\plot{no}{width=\textwidth}{Near-offset section.} \clearpage

\section{Common Reflection Angle Migration}
\inputdir{cram}

\plot{bpttidcrstk0}{width=\textwidth}{Image at 0-20 km interval.}
\multiplot{2}{bptttiocig0,bptttidcig0}{width=0.45\textwidth}
{Opening (a) and dip (b) angle gathers at 10 km.} \clearpage

\plot{bpttidcrstk1}{width=\textwidth}{Image at 20-40 km interval.}
\multiplot{2}{bptttiocig1,bptttidcig1}{width=0.45\textwidth}
{Opening (a) and dip (b) angle gathers at 20 km.}
\multiplot{2}{bptttiocig2,bptttidcig2}{width=0.45\textwidth}
{Opening (a) and dip (b) angle gathers at 30 km.} \clearpage

\plot{bpttidcrstk2}{width=\textwidth}{Image at 40-60 km interval.}
\multiplot{2}{bptttiocig3,bptttidcig3}{width=0.45\textwidth}
{Opening (a) and dip (b) angle gathers at 40 km.}
\multiplot{2}{bptttiocig4,bptttidcig4}{width=0.45\textwidth}
{Opening (a) and dip (b) angle gathers at 50 km.} \clearpage

\plot{bpttidcrstk3}{width=\textwidth}{Image at 60-78.75 km interval.}
\multiplot{2}{bptttiocig5,bptttidcig5}{width=0.45\textwidth}
{Opening (a) and dip (b) angle gathers at 60 km.}
\multiplot{2}{bptttiocig6,bptttidcig6}{width=0.45\textwidth}
{Opening (a) and dip (b) angle gathers at 70 km.} \clearpage
