\author{Johan Jensen}
%%%%%%%%%%%%%%%%%%%%%
\title{Homework 1}

\begin{abstract}
  This homework has three parts. 
  \begin{enumerate}
  \item Theoretical questions and computations related to the performance of 
  sorting algorithms.
  \item Analyzing digital elevation data from the Austin area. You will apply 
  histogram equalization to enhance the image.
  \item Analyzing seismic reflection data. You will apply an amplitude gain 
  correction to enhance the image.
  \end{enumerate}
\end{abstract}

\section{Prerequisites}

Completing the computational part of this homework assignment requires
\begin{itemize}
\item \texttt{Madagascar} software environment available from \\
\url{http://www.ahay.org/}
\item \LaTeX\ environment with \texttt{SEGTeX} available from \\ 
\url{http://www.ahay.org/wiki/SEGTeX}
\end{itemize}
To do the assignment on your personal computer, you need to install
the required environments. Please ask for help if you don't know where
to start.

The homework code is available from the \texttt{Madagascar} repository
by running
\begin{verbatim}
svn co https://rsf.svn.sourceforge.net/svnroot/rsf/trunk/book/geo391/hw1
\end{verbatim}

You can either write your answers on paper or edit them in the file
\verb#hw1/paper.tex#. Please show all the mathematical
derivations that you perform.

\newpage

\section{Sorting algorithms}
\inputdir{sort}  

\begin{itemize}
\item The cost of the \texttt{quantile} algorithm as 
a function of the data size $N$ can be estimated using 
the recursive relationship
\begin{equation}
\label{eq:cost}
P(N)=P(N/2)+C\,N\;,
\end{equation}
where $C$ is a constant. The recursion starts with $P(1)=0$. Derive an
explicit non-recursive formula for $P(N)$.
\item The rest of the homework involves computational assignments. 
      The first assignment involves sorting algorithms.

\lstset{language=c,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single]{sort/sort.c}

\newpage

\begin{enumerate}          
\item Change directory to \verb#geo391/hw1/sort#
\item Run
\begin{verbatim}
scons movie.vpl
\end{verbatim}
and observe a movie illustrating the slow data sorting algorithm. The
algorithm is implemented in the \texttt{slow\_sort} function in the
file \texttt{sort.c}.

\item Run
\begin{verbatim}
scons view
\end{verbatim}
to compute the cost of slow sorting experimentally. The output is
shown in Figure~\ref{fig:cost}.

\plot{cost}{width=\textwidth}{Experimental cost of slow sorting. 
The logarithm of the cost is shown against the logarithm of the data size.}

If we approximate the cost as $P(N)=N^\epsilon$, what is the value of
$\epsilon$ observed in the picture?

\item Open the file \texttt{sort.c} in a text editor and edit it to 
fix the specified line in the \texttt{quick\_sort} function.

\item Open the file \texttt{SConstruct} in a text editor and uncomment 
      the specified line.

\item Rerun
\begin{verbatim}
scons movie.vpl
\end{verbatim}
to observe a change in the sorting movie. Debug your changes to the
program if necessary.

\item Run
\begin{verbatim}
scons view
\end{verbatim}
to observe the change in the algorithm cost. What is the new
experimental value of $\epsilon$?

\item \textbf{EXTRA CREDIT} for improving the speed of the 
\texttt{quick\_sort} algorithm even further.
\end{enumerate}
\end{itemize}   

\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single]{sort/SConstruct}

\section{Histogram equalization}
\inputdir{dem}

\sideplot{byte}{width=\textwidth}{Digital elevation map of the 
west Austin area.}

Figure~\ref{fig:byte} shows a digital elevation map of the west Austin
area. Start by reproducing this figure on your screen.
\begin{enumerate}
\item Change directory to \verb#geo391/hw1/dem#
\item Run
\begin{verbatim}
scons byte.view
\end{verbatim}
\item Examine the file \texttt{byte.rsf} which refers to the 
byte (unsigned character) numbers which get displayed on the screen.
\begin{enumerate}
\item Open \texttt{byte.rsf} with a text editor to check its contents.
\item Run
\begin{verbatim}
sfin byte.rsf
\end{verbatim}
to check the data size and format.
\item Run
\begin{verbatim}
sfattr < byte.rsf
\end{verbatim}
to check data attributes. What is the maximum and minimum value? What
is the mean value? For an explanation of different attributes,
run \texttt{sfattr} without input.
\end{enumerate}
\end{enumerate}

Each image has a certain distribution of values (a histogram). The
histogram for the west Austin elevation map is shown in
Figure~\ref{fig:hist}. Notice the digitization artifacts. When
different values in a histogram are not uniformly distributed, the
image can have low contrast. One way of improving the contrast is
\emph{histogram equalization}.

\sideplot{hist}{width=\textwidth}{Normalized histogram (solid line)
  and cumulative histogram (dashed line) of the digital elevation
  data.}

Let $f(x,y)$ be the original image. The equalized image will be
$F(x,y)$. Let $h(f)$ be the histogram (probability distribution) of
the original image values. Let $H(F)$ be the histogram of the modified
image. The mapping of probabilities suggests
\begin{equation}
\label{eq:prob}
H(F)\,dF = h(f)\,df
\end{equation}
or, if we want the modified histogram
to be uniform, 
\begin{equation}
\label{eq:dif}
\frac{d F}{d f} = C\,h(f)\,
\end{equation}
where C is a constant. 
Solving equation~\ref{eq:dif}, we obtain the mapping
\begin{equation}
\label{eq:int}
F(f) = f_0 + C\,\int\limits_{f_0}^f h(\phi)\,d\phi\;,
\end{equation}
where $f_0$ is the minimum value of $f$.

The algorithm of histogram equalization consists of the following
three steps:
\begin{enumerate}
\item Taking an input image $f(x,y)$, compute its histogram $h(f)$.
\item Compute the cumulative histogram $F(f)$ according to 
equation~(\ref{eq:int}). Choose an appropriate $C$ so that the range
of $F$ is the same as the range of $f$.
\item Map every pixel $f(x,y)$ to the corresponding $F(x,y)$.
\end{enumerate}

Your task is:
\begin{enumerate}
\item Among the \texttt{Madagascar} programs, find a program that 
implements histogram equalization. \textbf{Hint:} you may find
the \texttt{sfdoc} utility useful.
\item Edit the \texttt{SConstruct} file to add histogram equalization. 
Create a new figure and compare it with Figure~\ref{fig:byte}.
\item Check the effect of equalization by recomputing the histogram 
in Figure~\ref{fig:hist} with equalized data. Run
\begin{verbatim}
scons hist.view
\end{verbatim}
to display the figure on your screen.`
\item \textbf{EXTRA CREDIT} for implementing the histogram
equalization algorithm independently.
\end{enumerate}	

\newpage

\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single]{dem/SConstruct}

\section{Time-power amplitude-gain correction}
\inputdir{tpow}

Raw seismic reflection data may come in the form of shot gathers
$S(x,t)$, where $x$ is the offset (horizontal distance from the
receiver to the source) and $t$ is recording time. Raw data are
inconvenient for analysis because of rapid amplitude decay of seismic
waves. The decay can be compensated by multiplying the data by a gain
function. A commonly used function is a power of time. The corrected
gather is
\begin{equation}
\label{eq:tpow}
S_\alpha(x,t) = t^{\alpha}\,S(x,t)\;.
\end{equation}
The advantage of the time-power gain is its simplicity and the ability
to reverse it by multiplying the data by $t^{-\alpha}$. What value of
$\alpha$ should one use? Jon Claerbout argues in favor of $\alpha=2$:
one factor of $t$ comes from geometrical spreading and the other from
multiple scattering. Your task is to develop an algorithm for finding
a better value for each specific data set.

\plot{tpow}{width=0.9\textwidth}{Seismic shot record before and after
  time-power gain correction.}

Figure~\ref{fig:tpow} shows a seismic shot record before and after
applying the time-power gain~(\ref{eq:tpow}) with $\alpha=2$. Start
by reproducing this figure on your screen.

\begin{enumerate}
\item Change directory to \verb#geo391/hw1/tpow#
\item Run
\begin{verbatim}
scons tpow.view
\end{verbatim}
\item Edit the \texttt{SConstruct} file. Find where the value of
  $\alpha$ is specified in this file and try changing it to a
  different value. Run \texttt{scons tpow.view} again to check the result.
\item How should one measure if the distribution of amplitudes after
  the gain correction is uniform? Suggest a measure (an objective
  function) that would take $S_\alpha(x,t)$ and produce one number that
  measures uniformity.
\item By modifying the program \texttt{objective.c}, compute your objective 
  function for different values of $\alpha$ and display it in a figure. 
  Does the function have a unique minimum or maximum?

\lstset{language=c,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single]{tpow/objective.c}

\item Suggest an algorithm for finding an optimal value of $\alpha$ by
  minimizing or maximizing the objective function. Your algorithm
  should be able to find the optimal value without scanning all
  possible values. \textbf{Hint:} if the objective function is
  $f(\alpha)=F[S_\alpha(x,t)]$ and 
\begin{equation}
\label{alpha}                      
f(\alpha) \approx f(\alpha_0) + 
  f'(\alpha_0)\,(\alpha-\alpha_0) + \frac{f''(\alpha_0)}{2}\,(\alpha-\alpha_0)^2
\end{equation}
then what is the optimal $\alpha$?
\item \textbf{EXTRA CREDIT} for implementing your algorithm for an automatic 
estimation of $\alpha$ and testing it on the shot gather from
Figure~\ref{fig:tpow}.

\end{enumerate}

\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single]{tpow/SConstruct}

\newpage

\section{Completing the assignment}

\begin{enumerate}
\item Change directory to \verb#geo391/hw1#.
\item Edit the file \texttt{paper.tex} in your favorite editor and change the
first line to have your name instead of Jensen's.
\item Run
\begin{verbatim}
sftour scons lock
\end{verbatim}
and
\begin{verbatim}
scons pdf
\end{verbatim}
\item Submit your result (file \texttt{paper.pdf}) on paper or by
e-mail.
\end{enumerate}
