\section{Estimation of interval parameters}

Similarly to effective parameters, interval parameters can also be
regarded as data attributes obtained from local slopes. Unlike the \tx
domain, the \taup domain offers us several alternatives to
conventional \citet{dix:68} processing for retrieving interval values.
In this section, we analyze three different options for extracting
interval parameters in the \taup domain.

\tabl{velocities}{This table lists the inputs required to retrieve
both effective and interval VTI parameters as data attributes obtained
from local slopes. The zero-slope time $\tau_0$ function is always
required to map the attributes to their correct position in
time. Processing data in \taup domain offers two more alternatives
than conventional Dix's processing in \tx domain. Moreover, we observe
that Fowler's is the only set of equations that {does not} require an
explicit use of the curvature.}{
\begin{center}
\large
\begin{tabular}{ cr | c | c | c | c | c | c |}
\cline{3-8}
	\multicolumn{2}{c}{}	  & \multicolumn{1}{|c|}{$\tau_{0}$} & $R$ &  $Q$  & $R_{\tau}$  & $Q_{\tau}$ & $\tau_{0,\tau}$ \\
\hline
\multicolumn{2}{|c|}{Effective $V_N$ $V_H$ $\eta$} & \tick & \tick & \tick  &   &  &\\
\hline
\multicolumn{1}{|c|}{\multirow{3}{*}{{Interval $\hat{V}_N$ $\hat{V}_H$ $\hat{\eta}$}}}
& {{Dix}}        & \tick  & \tick & \tick & \tick  & \tick &   \\
\cline{2-8}
\multicolumn{1}{|c|}{} 
& {{Stripping}} & \tick &         &       & \tick  & \tick &   \\
\cline{2-8}
\multicolumn{1}{|c|}{} 
& {{Fowler}}    & \tick  &         &       &  \tick &           & \tick  \\
\hline
\end{tabular}
\end{center}
}
\input{dix2}

\subsection{Claerbout's straightedge method}
\cite{Claerbout.sep.14.13} suggested that interval velocity in an
isotropic (\mbox{$\hat{V}_N=\hat{V}_H$}) layered medium could be
estimated with a \textit{pen and a straightedge} by measuring the
offset difference $\Delta x$ between equal slope $p$ points on two
reflection events (\rFg{commonray}a).  The computation of $\Delta x$
is straightforward after a CMP gather has been transformed into the
\taup domain. In fact, the \taup transform naturally aligns seismic
events with equal slope along the same trace
(\rFg{commonray}b). Moreover, local slopes $R(\tau,p)=d\tau/dp$ are
related to the emerging offset $x=-R$  \citep{baan:719} , therefore
Claerbout's inversion formula can be expressed as
\begin{equation}
\hat{V}_N^{2}(\tau,p)=\dfrac{R_{\tau}}{p^2R_{\tau}-p},  \label{eqn:Clarebout}
\end{equation}
where $R_{\tau}=\partial R(\tau,p)/\partial \tau$ and $\hat{V}_N$ is the
NMO interval velocity that we map back to zero-slope time $\tau_0$ using the
isotropic velocity independent \taup NMO, as suggested by \cite{fomel:S139}. The details of the derivation are in appendix C.

The two methods discussed next can be thought of as two alternative
extensions for VTI media of the original Claerbout's straightedge
method.

\input{stripping2}
\input{fowler2}

\section{Flattening by predictive painting}

As shown in Table \ref{tbl:velocities}, Fowler's is the only set of
equations that do not require an explicit use of the curvature
$Q$. The dependence on the curvature is absorbed by the $\tau_0$
function. The other two sets of equations, Dix and stripping formulas,
as well as the equations for effective parameters, do need
curvature. The curvature computation can be problematic when the data
are contaminated by noise. This makes these three methods (effective,
stripping, and Dix) less practical when applied to real data with poor
SNR. However, Fowler's rules represent a way to circumvent the
problem. In fact, if we can find an algorithm that estimates the $\tau_0$
mapping function directly from the data, all the curvature issues
will get solved.

%So next the question is: is there exist such an algorithm?

\inputdir{flattening}
%\plot{flat-0-all}{width=0.8\textwidth}{Sythetic image form
%\cite{claerbout:2006} (a), local dip estimate (b), flattened image
%(c) using the predicted time of reflecting surfaces (d) obtained
%after painting of a single reference time-trace. This example is
%adapted from \cite{fomel:A25}.}

The desired algorithm exists and is known as seismic image flattening. 
\begin{comment}
Flattening is a common task in geophysical interpretation and is
typically used to identify stratigraphic features within a seismic
image. It attempts to reverse the effects of geologic processes by
removing all the complex structures (\rFg{flat-0-all}a) present in the
image to make these structural features flat (\rFg{flat-0-all}c). A
flattened image is typically created by shifting samples in the
original image up or down. In the recent past we can find in the
literature different approaches and solutions to make this process
fully automated. The methods proposed by \cite{lomask:P13},
\cite{parks:2010} and \cite{fomel:acc} use an estimate of local image
slopes to estimate the vertical shift field at each sample in the
image.
\end{comment}
The idea
of using local slopes for automatic flattening was introduced by
\cite{bienati} and \cite{lomask:P13}. 
Flattening by predictive painting (appendix A)
%\cite[]{fomel:A25}  
uses the local-slope field 
%(\rfg{flat-0-all}b) 
to construct a recursive prediction operator (equation \ref{eq:pred}) that spreads a traveltime reference trace in the image and predicts the reflecting surfaces 
%(\rfg{flat-0-all}d) 
which are then unwrapped until the image is flattened.
%(\rfg{flat-0-all} (c) ).

\inputdir{synth}

We propose bypassing the issue of estimating the zero-slope time $\tau
_{0}$ field by using the predictive painting \old{method}
\new{approach}. Let us discuss how it works on the previously shown
synthetic data in \rFg{dataPsynth}a.
%\citep{fomel:864,fomel:acc}. Predictive painting is a numerical
%algorithm for automatic spreading of information along seismic data,
%according to the local structure of seismic events. Structure is
%commonly characterized by dominant local slopes
%\citep{ClareboutESA}. Thus, the method first finds slope-based
%prediction operators of each trace in the data from its neighbor
%traces.
% (equation 5 in \citep{fomel:864}). 
%In the second step, prediction operators are used to spread
%information along the local data slopes. 
\rFg{dataPsynth}b shows local event slope $R$ measured from the data
using the PWD algorithm. \rFg{dataPsynth}c shows how predictive
painting spreads a zero-slope time $\tau _{0}$ reference trace along
local data slopes to predict the zero-slope time $\tau _{0}$ mapping
field and hence the geometry of the traveltime reflection curves along
\taup CMP gather. Because this procedure does not involve curvature
computations, it represents a much more robust way of obtaining the
$\tau _{0}$ field that is needed by the inversion formulas in
equations \ref{eqn:fowler1} and \ref{eqn:fowler2}. After $\tau _{0}$
has been found, we also have what we need to perform gather flattening
\citep{burnett:WB129,burnett:3710}.
% which is effectively an alternative velocity-independent procedure for accomplishing
%\taup moveout correction.  
Unshifting each trace (\rFg{dataPsynth}d) automatically flattens the
data, thus performing a velocity-independent \taup NMO correction. As
expected, all events are perfectly aligned, and the correction does
not suffer from instabilities of curvature estimation.  Moreover,
predictive painting is automatic and does not require any prior
assumptions about the moveout shape.
%These properties makes the procedure also suitable for other
%tasks, such as the inversion of interval parameters for SV-wave data
%proposed by \cite{stovas:2506}.

\plot*{dataPsynth}{width=\linewidth}{Synthetic  CMP $\tau$-$p$ transformed gather (a), estimated local slopes (b), zero slope time $\tau_0$ obtained by predictive painting (c), and the gather flattened (d). \label{fig:dataSYNTH}}

Now, given the slope field $R$ and its zero-slope time field $\tau
_{0}$, we retrieve interval parameters using equations
\ref{eqn:fowler1} and \ref{eqn:fowler2}. In \rFg{intPmasksynth}, the
estimated NMO (a) horizontal velocities (b) and the anellipticity (c)
parameter are mapped to the appropriate zero-slope time using the
painted zero-slope time $\tau _{0}$ field (\rFg{dataPsynth}c).
%Since the synthetic are noiseless, the results are similar and the
%following observations still hold for the other approaches that employ
%Dix inversion, stripping equation and also Fowler's equation in which
%$\tau_0$ comes from the velocity independent moveout \req{tau0mapping}.
The exact interval profiles (yellow lines) are recovered \new{nearly}
perfectly although the resolution slightly worsens with respect to the
effective profiles (\rFg{eff-Syn}). The main reason is the instability
of the additional numerical differentiation along the $\tau $
direction that all the approaches require.

\inputdir{synth} 
\plot{intPmasksynth}{width=\textwidth}{Fowler's equation based
  inversion to interval normal moveout (a) horizontal velocity (b) and
  anellipticity parameter $\eta $ (c).  The yellow lines represent the
  exact values used for generating the synthetic dataset in figure
  \ref{fig:dataSynth} (a). }
