\section{Introduction}
A key challenge for imaging in complex areas is accurate determination
of a velocity model in the area under investigation.  Migration
velocity analysis is based on the principle that image accuracy
indicators are optimized when data are correctly imaged.  A common
procedure for velocity analysis is to examine the alignment of images
created with multi-offset data.  An optimal choice of image analysis
can be done in the angle domain which is free of some complicated
artifacts present in surface offset gathers in complex areas
\cite[]{GEO69-02-05620575}.

\par Migration velocity analysis after migration by wavefield
extrapolation requires image decomposition in scattering angles
relative to reflector normals. Several methods have been proposed for
such decompositions \cite[]{GEO55-09-12231234, SEG-1999-08240827,
SEG-2000-08300833, GEO67-03-08830889, XieWu.adcig, GEO68-03-10651074,
SEG-2003-08890892, Fomel.seg.3dadcig, GEO69-05-12831298}.  These
procedures require decomposition of extrapolated wavefields in
variables that are related to the reflection angle.

A key component of such image decompositions is the imaging condition.
A careful implementation of the imaging condition preserves all
information necessary to decompose images in their angle-dependent
components.  The challenge is efficient and reliable construction of
these angle-dependent images for velocity or amplitude analysis.

In migration with wavefield extrapolation, a prestack imaging
condition based on spatial shifts of the source and receiver
wavefields allows for angle-decomposition
\cite[]{GEO67-03-08830889,SavaFomel.pag}.  Such formed angle-gathers
describe reflectivity as a function of reflection angles and are
powerful tools for migration velocity analysis (MVA) or amplitude
versus angle analysis (AVA).  However, due to the large expense of
space-time cross-correlations, especially in three dimensions, this
imaging methodology is not used routinely in data processing.

This paper presents a different form of imaging condition.  The key
idea of this new method is to use time-shifts instead of space-shifts
between wavefields computed from sources and receivers.  Similarly to
the space-shift imaging condition, an image is built by space-time
cross-correlations of subsurface wavefields, and multiple lags of the
time cross-correlation are preserved in the image.  Time-shifts have
physical meaning that can be related directly to reflection geometry,
similarly to the procedure used for space-shifts.  Furthermore,
time-shift imaging is cheaper to apply than space-shift imaging, and
thus it might alleviate some of the difficulties posed by costly
cross-correlations in 3D space-shift imaging condition.

The idea of a time-shift imaging condition is related to the idea of
depth focusing analysis \cite[]{SEG-1986-S7.6,
GEO57-12-16081622,GEO58-08-11481156, SEG-1995-0465,SEG-1996-0463}.
The main novelty of our approach is that we employ time-shifting to
construct angle-domain gathers for prestack depth imaging.

The time-shift imaging concept is applicable to Kirchhoff migration,
migration by wavefield extrapolation, or reverse-time migration.  We
present a theoretical analysis of this new imaging condition, followed
by a physical interpretation leading to angle-decomposition.  Finally,
we illustrate the method with images of the complex Sigsbee 2A dataset
\cite[]{SEG-2002-21222125}.

