\author{Team: Longhorns}
%%%%%%%%%%%%%%%%%%%%%%%%
\title{GEO 365N/384S Seismic Data Processing \\ Computational Assignment 4}

\begin{abstract}
  This assignment consists of four parts:
  \begin{enumerate}
  \item Processing Alaska data through surface-consistent amplitude correction, velocity analysis, and stack.
  \item Processing Viking Graben data through deconvolution and surface-consistent amplitude correction.
  \item Processing Teapot Dome data through surface-consistent amplitude correction.
  \item Processing your own data.
  \end{enumerate}
\end{abstract}

\section{Prerequisites}

Completing the computational part of this homework assignment requires
\begin{itemize}
\item \texttt{Madagascar} software environment available from \\
\url{http://www.ahay.org/}
\item \LaTeX\ environment with \texttt{SEG}\TeX\ available from \\ 
\url{http://www.ahay.org/wiki/SEGTeX}
\end{itemize}
To do the assignment on your personal computer, you need to install
the required environments. 

To setup the Madagascar environment in the JGB 3.216B computer lab, run the following commands:
\begin{verbatim}
module load madagascar-devel
source $RSFROOT/share/madagascar/etc/env.csh
setenv DATAPATH $HOME/data/
setenv RSFBOOK $HOME/data/book
setenv RSFFIGS $HOME/data/figs
\end{verbatim}
You can put these commands in your \verb+$HOME/.cshrc+ file to run them automatically at login time.

To setup the \LaTeX\ environment, run the following commands:
\begin{verbatim}
cd
git clone https://github.com/SEGTeX/texmf.git
texhash
\end{verbatim}
You only need to do it once.

The homework code is available from the class repository
by running
\begin{verbatim}
svn co https://github.com/TCCS-BEG/geo384s/trunk/hw4
\end{verbatim}
You can also download it from your team's private repository.

\section{Generating this document}

At any point of doing this computational assignment, you can
regenerate this document and display it on your screen.

\begin{enumerate}          
\item Change directory to \texttt{hw4}:
\begin{verbatim}
cd hw4
\end{verbatim}
\item Run
\begin{verbatim}
sftour scons lock
scons read &
\end{verbatim}
\end{enumerate}

As the first step, open \texttt{hw4/paper.tex} file in your favorite
editor and edit the first line to enter the name of your team. Then
run \texttt{scons read} again.

\section{Alaska data}
\inputdir{alaska}

In the first part of the assignment, we will pick up processing of the
Alaska line where he last left it, after the groundroll attenuation
step. Figure~\ref{fig:rshots} shows the shot gathers after groundroll
removal by the \emph{Emc-Hammer} team.

\plot{rshots}{width=0.7\textwidth}{Alaska shot gathers after groundroll removal by \emph{Emc-Hammer}.}

In this assignment, you will process the data further by applying
surface-consistent amplitude balancing, velocity analysis, and stack.

\begin{enumerate}

\item Change directory to \texttt{hw4/alaska}.
\item Run
\begin{verbatim}
scons -c
\end{verbatim}
to remove (clean) previously generated files.
\item Run
\begin{verbatim}
scons arms.view
\end{verbatim}
to display trace amplitudes for Alaska shot gathers (Figure~\ref{fig:arms}). The amplitudes show a strong offset trend. To display them after removing the trend (Figure~\ref{fig:arms2}), run
\begin{verbatim}
scons arms2.view
\end{verbatim}

\multiplot{2}{arms,arms2}{width=0.45\textwidth}{Trace amplitudes before (a) and after (b) removing the long-period offset trend.}

Do you notice stripes in the amplitude going in different directions?
These stripes might be caused by near-surface conditions in deploying
sources and receivers. The surface-consistent model \cite[]{GEO46-01-00170022} tries to explain
the trace amplitude as a function of source and receiver locations
$A(s,r)$ using a product of several factors (source, receiver, offset, and
midpoint):
\begin{equation}
\label{eq:sc}
A(s,r) \approx A_s(s)\,A_r(r)\,A_x(r-s)\,A_m\left(\frac{r+s}{2}\right)\;.
\end{equation}
By applying logarithm to the amplitude, we can turn the product in equation~(\ref{eq:sc}) into a sum
\begin{equation}
\label{eq:sclog}
\log\left[A(s,r)\right] \approx L_s(s) + L_r(r) + L_x(r-s) + L_m\left(\frac{r+s}{2}\right)\;.
\end{equation}
\item In its discrete version, equation~(\ref{eq:sclog}) represents a system of linear equations with given $\log\left[A(s,r)\right]$ and unknown factors $L_s$, $L_r$, $L_x$, and $L_m$.

Does this system have more equations than unknowns or more unknowns than equations?

\answer{% Fill your answer here
}

Does this system have a unique solution?

\answer{% Fill your answer here
}

We will attempt to solve the system represented by
equation~(\ref{eq:sclog}) using the the least-squares method and the
iterative conjugate-gradient algorithm \cite[]{hestenes,fletcher}. The
algorithm requires implementing the linear operator and its
adjoint. The implementation of the linear operator from
equation~(\ref{eq:sclog}) is provided in
the \texttt{surface-consistent.c} program.

\lstset{language=c,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single,title=alaska/surface-consistent.c]{alaska/surface-consistent.c}

To run the iterative inversion and to see the result of matching the
data (logarithm of the amplitude), run
\begin{verbatim}
scons scarms.view
\end{verbatim}
Does it manage to predict the stripes seen in Figure~\ref{fig:arms2}?

\multiplot{2}{scarms,adiff}{width=0.45\textwidth}{(a) Estimated surface-consistent trace amplitudes. (b) Difference with Figure~\ref{fig:arms2}.}

The individual factors (composing the estimated model vector) are shown in Figure~\ref{fig:shot,receiver,offset,cmp}.

\multiplot{4}{shot,receiver,offset,cmp}{width=0.45\textwidth}{Estimated surface-consistent factors.}

To display the shot gathers after the surface-consistent amplitude correction (Figure~\ref{fig:ashots}), run
\begin{verbatim}
scons ashots.view
\end{verbatim}
To compare the output with the input, run
\begin{verbatim}
sfpen Fig/rshots.vpl Fig/ashots.vpl
\end{verbatim}

\plot{ashots}{width=0.7\textwidth}{Alaska shot gathers after surface-consistent amplitude balancing.}

\item After the correction, we can proceed with the data processing sequence by doing the familiar steps of sorting into CMP gathers, velocity analysis, NMO, and stack. Run
\begin{verbatim}
scons stack0.view
\end{verbatim}
to go through these steps. The resultant stack (Figure~\ref{fig:stack0}) appears to contain spiky noise. We can try to remove it by median filtering (Figure~\ref{fig:stack1})
\begin{verbatim}
scons stack1.view
\end{verbatim}

\multiplot{2}{stack0,stack1}{width=0.8\textwidth}{(a) First attempt at NMO stack. (b) NMO stack after despiking by median filtering.}

The NMO velocity extracted by the automatic picker is shown in
Figure~\ref{fig:vpicks}, and the velocity analysis and NMO applied to
an individual CMP gather are shown in Figure~\ref{fig:nmo2}, displayed with
\begin{verbatim}
scons nmo2.view
\end{verbatim}

\plot{vpicks}{width=0.8\textwidth}{NMO velocity estimated by the automatic picker.}
\plot{nmo2}{width=0.8\textwidth}{Velocity analysis and NMO applied to a selected CMP gather.}

\item Your task:
\begin{enumerate}
\item By modifying the \texttt{SConstruct} file, extract velocity analysis and NMO from several other CMP gathers and examine the results.
\item Try to improve the quality of the stack. To achieve a better result, you are free to use any of the techniques that you learned previously.
\end{enumerate}

\end{enumerate}


\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single,title=alaska/SConstruct]{alaska/SConstruct}

\section{Viking Graben data}
\inputdir{viking}

In the next part of the assignment, we will take a step back in
processing the Viking Graben dataset and use the original unprocessed data. 

\begin{enumerate}

\item Change directory to \texttt{hw4/viking}.
\item Run
\begin{verbatim}
scons -c
\end{verbatim}
to remove (clean) previously generated files.
\item In addition to seismic data, Mobil Oil provided a far-field wavelet recorded in the water. To display this wavelet, run
\begin{verbatim}
scons wavelet.view
\end{verbatim}
As typical for marine data, the wavelet, instead of being a perfect
impulse, contains additional components (bubble pulse, ghost
reflection, etc.) The wavelet spectrum, displayed with
\begin{verbatim}
scons spectrum.view
\end{verbatim}
shows frequencies missing from the ideal flat response.

\multiplot{2}{wavelet,spectrum}{width=0.45\textwidth}{Far-field wavelet (a) and its Fourier spectrum (b).}

\item The task of \emph{deconvolution} is to collapse the wavelet to
something closer to a pulse by convolving it with a
filter \cite[]{DEC00-00-04820482,DED00-00-07260726}. The filter
estimation is another classic application of the least-squares
inversion approach. The \emph{prediction-error filter} (PEF) is
defined by auto-regression: the filter coefficients represent weights,
which combine with shifted version of the input signal to predict the
signal itself:
\begin{equation}
\label{eq:pef}
-d_t \approx \sum_{i=1}^{N}\,f_i\,d_{t-i}\;.        
\end{equation}
In equation~(\ref{eq:pef}), $d_t$ is the given data, and $f_i$ for
$i=1,2,\ldots,N$ are free coefficients of the prediction-error filter. 

As in the surface-consistent problem, the least-squares solution can
be found using a generic conjugate-gradient program, which simply
requires an implementation of the convolution operator in the
right-hand side of equation~(\ref{eq:pef}) and its adjoint. An
implementation is provided in the \texttt{convolve.c} program.

\lstset{language=c,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single,title=viking/convolve.c]{viking/convolve.c}

To test the program using the generic dot-product test, you can run
\begin{verbatim}
scons filter0.rsf wavelet4.rsf
scons convolve.exe
sfdottest ./convolve.exe nf=100 mod=filter0.rsf dat=wavelet4.rsf
\end{verbatim}
The test should show pairs of numbers matching to several significant digits.

To run the least-squares inversion and to estimate the filter, run
\begin{verbatim}
scons filter.view
\end{verbatim}

\multiplot{2}{filter,wdecon}{width=0.45\textwidth}{Estimated prediction-error filter (a) and the result of wavelet deconvolution (b).}

Although deconvolution applied to the wavelet
(Figure~\ref{fig:wdecon}) does not produce a perfect spike, it
improves the spikiness of the wavelet and its spectral content.

\item Now we can apply the estimated filter to the data. Run
\begin{verbatim}
scons viking1000.view
\end{verbatim}
to extract and display the first 1,000 traces from the Viking Graben
dataset (Figure~\ref{fig:viking1000}.) Run
\begin{verbatim}
scons decon1000.view
\end{verbatim}
to display the result of deconvolution (Figure~\ref{fig:decon1000}.)

\plot{viking1000}{width=0.8\textwidth}{First 1,000 traces from the Viking Graben dataset.}

\plot{decon1000}{width=0.8\textwidth}{Viking Graben traces after deconvolution.}

To apply deconvolution to the whole dataset, run
\begin{verbatim}
scons decon.rsf
\end{verbatim}

\item (\textbf{EXTRA CREDIT}) For an extra credit, you can write a program or a script, which takes the data as the main input and the filter as an auxiliary input and applies deconvolution. With your new program, you can replace the lines
\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\begin{lstlisting}
# Process all traces
Flow('decon',['filter','viking',convolve],
     '''
     ./\${SOURCES[2]} data=i\${SOURCES[1]} adj=n | 
     add \${SOURCES[1]} scale=-1,1
     ''')
\end{lstlisting}
with something like
\begin{lstlisting}
# Process all traces
Flow('decon',['viking','filter',decon],
     '''
     ./\${SOURCES[2]} filter=\${SOURCES[1]}  
     ''',split=[2,'omp',[0]])
\end{lstlisting}
and process the data quickly in parallel.

\item After deconvolution, we can proceed to surface-consistent amplitude correction, similar to the processing applied previously to the Alaska data. The trace amplitudes are shown in Figure~\ref{fig:varms}. Their surface-consistent estimate, using only the shot term $L_s$ and the offset term $L_x$ is shown in Figure~\ref{fig:vscarms}. The difference is shown in Figure~\ref{fig:vadiff}. To display the latter figure, run
To apply deconvolution to the whole dataset, run
\begin{verbatim}
scons vadiff.view
\end{verbatim}
Can you notice remaining stripes in the difference? Which of the four terms in equation~(\ref{eq:sclog}) are the stripes coming from?

\answer{
}

\multiplot{3}{varms,vscarms,vadiff}{width=0.3\textwidth}{Trace amplitudes from the Viking Graben dataset. (a) Initial. (b) Estimated surface-consistent. (c) Difference.}

\item Your task: modify the \texttt{SConstruct} file to improve the result
by introducing additional terms to the fitting equation.

\end{enumerate}

\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single,title=viking/SConstruct]{viking/SConstruct}

\section{Teapot Dome data}
\inputdir{teapot}

Next, we return to processing the Teapot Dome data. Our task is to
estimate the surface-consistent amplitude normalization, similar to
how it was done with the two previous datasets.

\begin{enumerate}

\item Change directory to \texttt{hw4/teapot}.
\item Run
\begin{verbatim}
scons -c
\end{verbatim}
to remove (clean) previously generated files.
\item The process should look familiar by now. Figure~\ref{fig:tshot,treceiver} shows the two amplitude factors ($L_s$ and $L_r$) estimated by least-squares inversion.
\item To visualize how the amplitude correction depends on the surface coordinates, we can display it by interpolating to a regular grid from the shot locations. To display the result (Figure~\ref{fig:sint}), run
\begin{verbatim}
scons sint.view
\end{verbatim}
\item Your task: modify the \texttt{SConstruct} file to interpolate in receiver coordinates (\texttt{gx} and \texttt{gy}) instead of shot coordinates (\texttt{sx} and \texttt{sy}). Do you notice any interesting differences?

\end{enumerate}

\multiplot{2}{tshot,treceiver}{width=0.45\textwidth}{Estimated shot and receiver surface-consistent amplitude terms for the Teapot Dome dataset.}

\plot{sint}{width=0.8\textwidth}{Surface-consistent amplitude interpolated using the shot coordinates.}

\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single,title=teapot/SConstruct]{teapot/SConstruct}

\section{Your own data}

Finally, add results from analyzing the dataset you selected for your
course project. You can include the first steps of analyzing the data
acquisition geometry and/or any other processing steps that you have
taken.

\section{Saving your work}

It is important to save all files that you edit by hand (such
as \texttt{paper.tex} and \texttt{SConstruct}) in a version-control
system every time you make a revision.  The completed assignment is
due in two weeks and should be submitted through your team's private
GitHub repository.

\bibliographystyle{seg}
\bibliography{SEG,sc}


