\author{Sergey Fomel, Parvaneh Karimi, Alexander Klokov, and Xiaolei Song}
\title{BP 1997 2.5-D model}

\maketitle

\begin{abstract}
John Etgen describes this model as ``The Carpathians thrusting over the North Sea''. It was created for the 1998 SEG talk \cite[]{SEG-1998-0066} and is also mentioned by \cite{GEO65-03-09430950}. It is provided courtesy of BP.
\end{abstract}

\section{Model}
\inputdir{model}

\plot{model}{width=\textwidth}{Velocity model (section from 2.5-D).} \clearpage

\plot{shots}{width=\textwidth}{Shot gathers from
  finite-difference modeling.} \clearpage
\plot{zo}{width=\textwidth}{Zero-offset data from
  finite-difference modeling.} \clearpage

\plot{refl}{width=\textwidth}{Exploding reflector.} \clearpage
\plot{twexp}{width=\textwidth}{High-resolution zero-offset data modeled from exploding-reflector modeling.} \clearpage


\section{Lowrank RTM}
\inputdir{lowrank}

\plot{zomigt}{width=\textwidth}{Zero-offset RTM image by the lowrank method using high-resolution data.} \clearpage

\section{FFD RTM}
\inputdir{ffd}

%\plot{rtm}{width=\textwidth}{Prestack RTM image by Fourier finite-differences.} \clearpage
\plot{rtmth}{width=\textwidth}{Prestack RTM image by Fourier finite-differences.} \clearpage

\section{Multi-arrival Kirchhoff migration}
\inputdir{kirMulti}

\plot{twohalf-timage}{width=\textwidth}{Prestack multi-arrival Kirchhoff migration.} \clearpage
\plot{fig8800}{width=\textwidth}{Common-image gather in (a) the dip-angle domain and 
(b) the scattering-angle domain from position 8.8~km.} \clearpage
\plot{fig15500}{width=\textwidth}{Common-image gather in (a) the dip-angle domain and 
(b) the scattering-angle domain from position 15.5~km.} \clearpage

\section{One-way wave extrapolation}
\inputdir{oway}

\plot{mig}{width=\textwidth}{Zero-offset image by one-way wave-equation migration using extended split-step approximation.} \clearpage

\bibliographystyle{seg}
\bibliography{SEG}
