\section{Discussion}

For the general anisotropic media, qP- and qS-wave modes are intrinsically coupled.
The elastic wave equation must be solved at once to get correct kinematics and amplitudes
for all modes. The scalar wavefields, however, are widely used with the help of
wave mode separation or by using approximate equations derived from the elastic wave equation 
for many applications such as seismic imaging.
As demonstrated in the theoretical parts, the pseudo-pure-mode wave equation is derived
from the elastic wave equation through a similarity transformation to the Christoffel
equation in the wavenumber domain. The components of the transformed wavefield essentially
represent the spatial derivatives of the displacement wavefield components.
This transformation preserves the kinematics of wave propagation but inevitablely
changes the phases and amplitudes for qP- and qS-waves as the 
elastic wave mode separation procedure using divergence-like and curl-like operations
\cite[]{dellinger.etgen:1990,yan.sava:2009,zhang.mcmechan:2010}.
The filtering step to correct the projection deviation is indispensable for complete
removing the residual qS-waves from the extrapolated pseudo-pure-mode qP-wave fields.
This procedure does not change the phases and amplitudes of the scalar qP-waves because
the deviation operator is computed using the normalized wave and polarization vectors.

In fact, it is not even clear what the correct amplitudes should be for "scalar anisotropy".
Like the anisotropic pseudo-analytic methods \cite[]{etgen:2009,fomel:2013,zhan:2012,song:2013},
the pseudo-pure-mode wave equation may distort the reflection, transmission and conversion
coefficients of the elastic wavefields when there are high-frequency perturbations in the velocity model.
Therefore, the converted qP-waves remaining in the separated qP-wave fileds only have reliable kinematics. 
What happens to the qP-wave's amplitudes and how to
make use of the converted qP-waves (for seismic imaging) 
on the base of the pseudo-pure-mode qP-wave equation need further investigation in our future research.
