%\shortnote

\lefthead{Fomel \& Claerbout}
\righthead{Implicit extrapolation}

\published{SEP report, 95, 43-60 (1997)}
\title{Exploring three-dimensional implicit wavefield extrapolation
  with the helix transform}

\email{sergey@sep.stanford.edu, jon@sep.stanford.edu}
%\keywords{depth migration, post-stack, finite-difference, velocity continuation, helix}

\author{Sergey Fomel and Jon F. Claerbout}

\maketitle

\begin{abstract}
  Implicit extrapolation is an efficient and unconditionally stable
  method of wavefield continuation. Unfortunately, implicit wave
  extrapolation in three dimensions requires an expensive solution of
  a large system of linear equations. However, by mapping the
  computational domain into one dimension via the helix transform, we
  show that the matrix inversion problem can be recast in terms of an
  efficient recursive filtering. Apart from the boundary conditions,
  the solution is exact in the case of constant coefficients (that is,
  a laterally homogeneous velocity.) We illustrate this fact with an
  example of three-dimensional velocity continuation and discuss
  possible ways of attacking the problem of lateral variations.
\end{abstract}

\section{Introduction}

Implicit finite-difference wavefield extrapolation played an
exceptionally important role in the early development of seismic
migration methods. Using limited-degree approximations to the one-way
wave equation, implicit schemes have provided efficient and
unconditionally stable numerical wave extrapolation operators
\cite[]{Godfrey.sep.16.83,Claerbout.blackwell.85}. Unfortunately, the
advantages of \emph{implicit} methods were lost with the development
of three-dimensional seismic exploration. While the cost of 2-D
implicit extrapolation is linearly proportional to the mesh size, the
same approach, applied in the 3-D case, leads to a nonlinear
computational complexity. Primarily for this reason, implicit
extrapolators were replaced in practice by \emph{explicit} ones,
capable of maintaining linear complexity in all dimensions. A number
of computational tricks \cite[]{GEO56-11-17701777} allow the commonly
used explicit schemes to behave stably in practical cases.  However,
their stability is not unconditional and may break in unusual
situations \cite[]{SEG-1994-1266}.
\par
In this paper, we present an approach to three-dimensional
extrapolation, based on the helix transform of multidimensional
filters to one dimension \cite[]{Claerbout.gem.97}. The traditional
approach involves an inversion of a banded matrix (tridiagonal in the
2-D case and blocked-tridiagonal in the 3-D case). With the help of
the helix transform, we can recast this problem in terms of inverse
recursive filtering.  The coefficients of two-dimensional filters on a
helix are obtained by one-dimensional spectral factorization methods.
As a result, the complexity of three-dimensional implicit
extrapolation is reduced to a linear function of the computational
mesh size. This approach doesn't provide an exact solution in the
presence of lateral velocity variations. Nevertheless, it can be used
for preconditioning iterative methods, such as those described by
\cite{Nichols.sep.70.31}.  In this paper, we demonstrate the
feasibility of 3-D implicit extrapolation on the example of laterally
invariant velocity continuation and, in the final part, discuss
possible strategies for solving the problem of lateral variations.
\par
The main application of finite-difference wave extrapolation is
\emph{post-stack} depth migration. An application of similar methods
for \emph{prestack} common-shot migration is constrained by the
limited aperture of commonly used seismic acquisition patterns.
Recently developed acquisition methods, such as the vertical cable
technique \cite[]{SEG-1993-1376}, open up new possibilities for 3-D wave
extrapolation applications. An alternative approach is common-azimuth
migration \cite[]{Biondi.sep.80.109,Biondi.sep.93.1}. Other interesting
applications include finite-difference data extrapolation in offset
\cite[]{Fomel.sep.84.179}, migration velocity \cite[]{Fomel.sep.92.159},
and anisotropy \cite[]{Alkhalifah.sep.94.tariq3}.

\section{Implicit versus Explicit extrapolation}

The difference between implicit and explicit extrapolation is best
understood through an example. Following \cite{Claerbout.blackwell.85},
let us consider, for instance, the diffusion (heat conduction) equation
of the form
\begin{equation}
  {\frac{\partial T}{\partial t}} = {a (x)\,{\frac{\partial^2 T}{\partial x^2}}}\;.
\label{eqn:heat}
\end{equation}
Here $t$ denotes time, $x$ is the space coordinate, $T (x,t)$ is the
temperature, and $a$ is the heat conductivity coefficient.
Equation (\ref{eqn:heat}) forms a well-posed boundary-value problem if
supplied with the initial condition
\begin{equation}
  \label{eqn:heatinit}
  \left.T\right|_{t=0} = T_0 (x)
\end{equation}
and the appropriate boundary conditions. Our task is to build a
digital filter, which transforms a gridded temperature $T$ from one
time level to another.
\par
It helps to note that when the conductivity coefficient $a$ is
constant and the space domain of the problem is infinite (or periodic)
in $x$, the problem can be solved in the wavenumber domain. Indeed,
after the Fourier transform over the variable $x$, equation
(\ref{eqn:heat}) transforms to the ordinary differential equation
\begin{equation}
  {\frac{d \hat{T}}{d t}} = {- a k^2\, \hat{T}}\;,
  \label{eqn:heatk}	
\end{equation}
which has the explicit analytical solution
\begin{equation}
  \label{eqn:heatsol}
  \hat{T} (k,t) = \hat{T}_0 (k) e^{- a k^2 t}\;,
\end{equation}
where $\hat{T}$ denotes the Fourier transform of $T$, and $k$ stands
for the wavenumber. Therefore, the desired filter in the
wavenumber domain has the form
\begin{equation}
  \label{eqn:heatf}
  H (k) = e^{- a k^2}\;,
\end{equation}
where for simplicity the coefficient $a$ is normalized for the time
step $\triangle t$ equal to $1$.
\par
Returning now to the time-and-space domain, we can approach the filter
construction problem by approximating the space-domain response of
filter (\ref{eqn:heatf}) in terms of the differential operators
$\frac{\partial^2}{\partial x^2} = - k^2$, which can be approximated
by finite differences. An \emph{explicit} approach would amount to
constructing a series expansion of the form
\begin{equation}
  \label{eqn:heatexpl}
  H_{\mbox{ex}} (k) \approx a_0 + a_1 k^2 + a_2 k^4 + \ldots\;,
\end{equation}
and selecting the coefficients $a_j$ to approximate equation
(\ref{eqn:heatf}). For example, the three-term Taylor series expansion
around the zero wavenumber yields
\begin{equation}
  \label{eqn:heattayl}
  H_{\mbox{ex}} (k) = 1 - a\,{k^2}  +
  {\frac{{{a }^2}\,{k^4}}{2}} \;.
\end{equation}
The error of approximation (\ref{eqn:heattayl}) as a function of $k$
for two different values of $a$ is shown in the left plot of Figure
\ref{fig:exerror,imerror}.

\inputdir{Sage}

\multiplot{2}{exerror,imerror}{width=2.8in}{Errors of second-order
  explicit (a) and implicit (b) approximations for the heat extrapolation.}
\par
An \emph{implicit} approach also approximates the ideal filter
(\ref{eqn:heatf}), but with a rational approximation of the form
\begin{equation}
  \label{eqn:heatpade}
  H_{\mbox{im}} (k) \approx \frac{b_0 + b_1 k^2 + b_2 k^4 + \ldots}
  {1 + c_1 k^2 + c_2 k^4 + \ldots}
\;.
\end{equation}
One way of selecting the coefficients $b_i$ and $c_i$ is to apply an
appropriate Pad\'{e} approximation \cite[]{pade}\footnote{If the
  denominator and the numerator have the same order, Pad\'{e}
  approximants are equivalent to the corresponding continuous
  fraction expansions.}.  For example the $[2/2]$ Pad\'{e}
approximation is
\begin{equation}
  \label{eqn:heatcrank}
  H_{\mbox{im}} (k) =
  \frac{1 - \frac{a}{2}\,k^2}{1 + \frac{a}{2}\,k^2}
  \;.
\end{equation}
This approximation corresponds to the famous Crank-Nicolson implicit
method \cite[]{cn}. The error of approximation (\ref{eqn:heatcrank}) as
a function of $k$ for different values of $a$ is shown in the right
plot of Figure \ref{fig:exerror,imerror}. Not only is it significantly smaller
than the error of the same-order explicit approximation, but it also
has a negative sign. It means that the high-frequency numerical noise
gets suppressed rather than amplified. In practice, this property
translates into a stable numerical extrapolation.
\par
The second derivative operator $-k^2$ can be approximated in practice
by a digital filter. The most commonly used filter has the
$Z$-transform $D_2 (Z) = -Z^{-1} + 2 - Z$, and the Fourier transform
\begin{equation}
  \label{eqn:d2k}
  D_2 (k) = e^{-ik} - 2 + e^{-ik} = 2 (\cos{k} - 1) = -4
  \sin^2{\frac{k}{2}}\;.
\end{equation}
Formula (\ref{eqn:d2k}) approximates $-k^2$ well only for small
wavenumbers $k$. As shown in Appendix A, the implicit scheme allows
the accuracy of the second-derivative filter to be significantly
improved by a variation of the ``1/6-th trick''
\cite[]{Claerbout.blackwell.85}. The final form of the implicit
extrapolation filter is
\begin{equation}
  \label{eqn:heatfk}
   H_{\mbox{im}} (k) =
   \frac{1 + \frac{a+\beta}{2}\,D_2 (k)}{1 - \frac{a-\beta}{2}\,D_2 (k)}
   \;,
\end{equation}
where $\beta$ is a numerical constant, found in Appendix A.

\inputdir{heat}

\plot{heat}{width=6in,height=2.5in}{Heat extrapolation with explicit
  and implicit finite-different schemes. Explicit extrapolation
  appears stable for $a=2/3$ (left plot) and unstable for $a=4/3$
  (middle plot). Implicit interpolation is stable even for larger
  values of $a$ (right plot).}
\par
A numerical 1-D example is shown in Figure \ref{fig:heat}. The initial
temperature distribution is given by a step function. The
discontinuity at the step gets smoothed with time by the heat
diffusion. The left plot shows the result of an explicit extrapolation
with $a=2/3$, which appears stable. The middle plot is an explicit
extrapolation with $a=4/3$, which shows a terribly unstable behavior:
the high-frequency numerical noise is amplified and dominates the
solution. The right plot shows a stable (though not perfectly
accurate) extrapolation with the implicit scheme for the larger value of
$a=2$.
\par
The difference in stability between explicit and implicit schemes is
even more pronounced in the case of \emph{wave extrapolation}. For
example, let us consider the ideal depth extrapolation filter in the
form of the phase-shift operator
\cite[]{GEO43-07-13421351,Claerbout.blackwell.85}
\begin{equation}
  \label{eqn:gazdag}
  W (k) = e^{i \sqrt{a^2 - k^2}}\;,
\end{equation}
where $a = \omega / v$, $\omega$ is the time frequency, and $v$ is the
seismic velocity (which may vary spatially); we assume for simplicity
that both the depth step $\triangle z$ and the space sampling
$\triangle x$ are normalized to $1$.  A simple implicit approximation
to filter (\ref{eqn:gazdag}) is
\begin{equation}
  \label{eqn:wave45}
   W_{\mbox{im}} (k) = e^{i a}\,
   \frac{1 -4 a^2 + i a\,k^2}{1 - 4 a^2 - i a\,k^2} = e^{i \phi}\;,
\end{equation}
where $\phi = a - 2 \arctan{\frac{a\,k^2}{4 a^2-1}}$. We can see
that approximation (\ref{eqn:wave45}) is again a pure phase shift
operator, only with a slightly different phase. For that reason, the
operator is unconditionally stable for all values of $a$: the total
wave energy from one depth level to another is preserved. Operator
(\ref{eqn:gazdag}) corresponds to the Crank-Nicolson scheme for the
45-degree one-way wave equation \cite[]{Claerbout.blackwell.85}. Its
phase error as a function of the dip angle $\theta =
\arcsin{\frac{k}{a}}$ for different values of $a$ is shown in Figure
\ref{fig:phase}.

\inputdir{Sage}

\sideplot{phase}{width=2.5in}{The phase error of the
  implicit depth extrapolation with the Crank-Nicolson method.}
\par
The unconditional stability property is not achievable with the
explicit approach, though it is possible to increase the stability of
explicit operators by using relatively long filters
\cite[]{GPR36-02-00990114,GEO56-11-17701777}.

\section{Spectral factorization and three-dimensional extrapolation}

In this section, we continue our review of extrapolation methods to
reveal the principal difficulties of three-dimensional extrapolation.
We then describe a new, helix-transform approach to this old and
fascinating problem.

\subsection{Inverse filter factorization}
The conventional way of applying implicit finite-difference schemes
reduces to solving a system of linear equations with a sparse matrix.
For example, to apply the scheme of equation (\ref{eqn:heatfk}), we
can put the filter denominator on the other side of the extrapolation
equation, writing it as
\begin{equation}
  \label{eqn:heattris}
  \left(\mathbf{I} - \frac{a-\beta}{2}\,\mathbf{D}_2\right) \mathbf{T}_{t+1} =
  \left(\mathbf{I} + \frac{a+\beta}{2}\,\mathbf{D}_2\right) \mathbf{T}_t\;,
\end{equation}
where $\mathbf{I}$ is the identity matrix, $\mathbf{D}$ is the convolution
matrix for filter (\ref{eqn:d2k}), and $\mathbf{T}_t$ is the vector of
temperature distribution at time level $t$. In the case of
two-dimensional extrapolation, the matrix on the left side of equation
(\ref{eqn:heattris}) takes the tridiagonal form
\begin{equation}\label{eqn:matrix}
  \mathbf{A} = \left(\mathbf{I} -c\,\mathbf{D}_2\right) =
  \left[\begin{array}{cccccc}
  1+2 c_{1}   & -c_{1}     &  0     & \cdots &        & 0      \\
  -c_{2}     & 1+2c_{2}   & -c_{2}     & 0      & \cdots &        \\
  0      & -c_{3}     & \cdots & \cdots &        & \cdots \\
  \cdots & 0      & \cdots &        &        &        \\
         & \cdots &        &        & \cdots & -c_{n-1}     \\
  0      &        &        &        & -c_{n}     & 1 + 2c_{n}
  \end{array}\right]\;,
\end{equation}
where $c = \frac{a-\beta}{2}$, and where, for simplicity, we assume
zero-slope boundary conditions. Like any positive-definite tridiagonal
matrix, matrix $\mathbf{A}$ can be inverted recursively by an $LU$
decomposition into two bidiagonal matrices. The cost of inversion is
directly proportional to the number of vector components.  The same
conclusion holds for the case of depth extrapolation [equation
(\ref{eqn:wave45})] with the substitution $c = \frac{\beta + i
  a}{1-4\,a^2}$.
\par
In the case of a laterally constant coefficient $a$, we can take a
different point of view on the tridiagonal matrix inversion. In this
case, the matrix $\mathbf{A}_2$ represents a convolution with a
symmetric three-point filter $1-c\,D_2(k)$. The $LU$
decomposition of such a matrix is precisely equivalent to filter
\emph{factorization} into the product of a causal minimum-phase filter
with its adjoint. This conclusion can be confirmed by the easily
verified equality
\begin{equation}\label{eqn:d2d1}
  1 + c (Z^{-1} - 2 + Z) = \frac{(1+b)^2}{4}\, \left(1 + \frac{1-b}{1+b} Z\right)
  \,\left(1 + \frac{1-b}{1+b} Z^{-1}\right)\;,
\end{equation}
where $b = \sqrt{1+ 4\,c}$. The inverse of the causal minimum-phase
filter $1 + \frac{1-b}{1+b} Z$ is a recursive inverse filter.
Correspondingly, the inverse of its adjoint pair, $1 + \frac{1-b}{1+b}
Z^{-1}$, is the same inverse filtering, performed in the adjoint mode
(backwards in space).  In the next subsection, we show how this
approach can be carried into three dimensions by applying the helix
transform.

\subsection{Helix and multidimensional deconvolution}
\inputdir{Math}

The major obstacle of applying an implicit extrapolation in three
dimensions is that the inverted matrix is no longer tridiagonal. If we
approximate the second derivative (Laplacian) on the 2-D plane with
the commonly used five-point filter $Z_x^{-1} + Z_y^{-1} -4 + Z_x +
Z_y$, then the matrix on the left side of equation
(\ref{eqn:heattris}), under the usual mapping of vectors from a
two-dimensional mesh to one dimension, takes the infamous
blocked-tridiagonal form \cite[]{birk}
\begin{equation}\label{eqn:matrix2}
  \tilde{\mathbf{A}} = \left(\mathbf{I} -c\,\mathbf{D}_2\right) =
  \left[\begin{array}{cccccc}
  \mathbf{A}_1   & -c_1 \,\mathbf{I}    &  0     & \cdots &        & 0      \\
  -c_2 \,\mathbf{I}    & \mathbf{A}_2   & -c_2 \,\mathbf{I}    & 0      & \cdots &        \\
  0      & -c_3 \,\mathbf{I}     & \cdots & \cdots &        & \cdots \\
  \cdots & 0      & \cdots &        &        &        \\
         & \cdots &        &        & \cdots & -c_{n-1} \,\mathbf{I}    \\
  0      &        &        &        & -c_{n} \,\mathbf{I}    &
  \mathbf{A}_n
  \end{array}\right]\;.
\end{equation}
Inspecting this form more closely, we see that the main diagonal of
$\tilde{\mathbf{A}}$, as well as the two offset diagonals formed by the
scaled identity matrices, remains continuous, while the second top and
bottom diagonals are broken. Therefore, even for constant $c$, the
inverted matrix does not have a simple convolutional structure, and
the cost of its inversion grows nonlinearly with the number of grid
points.
\par
A \emph{helix transform}, recently proposed by one of the authors
\cite[]{Claerbout.sep.95.jon1}, sheds new light on this old problem.
Let us assume that the extrapolation filter is applied by sliding it
along the $x$ direction in the $\{x,y\}$ plane. The diagonal
discontinuities in matrix $\tilde{\mathbf{A}}$ occur exactly in the
places where the forward leg of the filter slides outside the
computational domain. Let's imagine a situation, where the leg of the
filter that went to the end of the $x$ column, would immediately
appear at the beginning of the next column. This situation defines a
different mapping from two computational dimensions to the one
dimension of linear algebra. The mapping can be understood as the
helix transform, illustrated in Figure \ref{fig:helix1} and explained
in detail by \cite{Claerbout.sep.95.jon1}. According to this
transform, we replace the original two-dimensional filter with a long
one-dimensional filter. The new filter is partially filled with zero
values (corresponding to the back side of the helix), which can be
safely ignored in the convolutional computation.

\plot{helix}{width=5in,bb=210 155 630 390}{The helix transform of
  two-dimensional filters to one dimension. The two-dimensional filter
  in the left plot is equivalent to the one-dimensional filter in the
  right plot, assuming that a shifted periodic condition is imposed on
  one of the axes.}
\par
This is exactly the helix transform that is required to make all the
diagonals of matrix $\tilde{\mathbf{A}}$ continuous. In the case of
laterally invariant coefficients, the matrix becomes strictly Toeplitz
(having constant values along the diagonals) and represents a
one-dimensional convolution on the helix surface. Moreover, this
simplified matrix structure applies equally well to larger
second-derivative filters ( such as those described in Appendix B),
with the obvious increase of the number of Toeplitz diagonals.
Inverting matrix $\tilde{\mathbf{A}}$ becomes once again a simple
inverse filtering problem.  To decompose the 2-D filter into a pair
consisting of a causal minimum-phase filter and its adjoint, we can
apply spectral factorization methods from the 1-D filtering theory
\cite[]{Claerbout.blackwell.76,Claerbout.blackwell.92}, for example,
Kolmogorov's highly efficient method \cite[]{kolmog}. Thus, in the case
of a laterally invariant implicit extrapolation, matrix inversion
reduces to a simple and efficient recursive filtering, which we need
to run twice: first in the forward mode, and second in the adjoint
mode.

\inputdir{heat}

\plot{heat3d}{width=6in,height=2in}{Heat extrapolation in two
  dimensions, computed by an implicit scheme with helix recursive
  filtering. The left plot shows the input temperature distributions;
  the two other plots, the extrapolation result at different time
  steps.  The coefficient $a$ is 2.}
\par
Figure \ref{fig:heat3d} shows the result of applying the helix
transform to an implicit heat extrapolation of a two-dimensional
temperature distribution. The unconditional stability properties are
nicely preserved, which can be verified by examining the plot of
changes in the average temperature (Figure \ref{fig:heat-mean}).

\sideplot{heat-mean}{width=2.5in}{Demonstration of the stability of
  implicit extrapolation. The solid curve shows the normalized mean
  temperature, which remains nearly constant throughout the
  extrapolation time. The dashed curve shows the normalized maximum
  value, which exhibits the expected Gaussian shape.}
\par
In principle, we could also treat the case of a laterally invariant
coefficient with the help of the Fourier transform. Under what
circumstances does the helix approach have an advantage over Fourier
methods? One possible situation corresponds to a very large input data
size with a relatively small extrapolation filter. In this case, the
$O(N log N)$ cost of the fast Fourier transform is comparable with the
$O(N_f N)$ cost of the space-domain deconvolution (where $N$
corresponds to the data size, and $N_f$ is the filter size). Another
situation is when the boundary conditions of the problem have an
essential lateral variation. The latter case may occur in applications
of velocity continuation, which we discuss in the next section.  Later
in this paper, we return to the discussion of problems associated with
lateral variations.

\section{Three-dimensional implicit velocity continuation}
\inputdir{vc3}

Velocity continuation is a process of navigating in the migration
velocity space, applicable for time migration, residual migration, and
migration velocity analysis \cite[]{Fomel.sep.92.159}. In the
zero-offset (post-stack) case, the velocity continuation process is
described by the simple partial differential equation
\cite[]{Claerbout.sep.48.79,me}
\begin{equation}
  {\frac{\partial^2 P}{\partial v\,\partial t}} +
  {v\,t\,\left({\frac{\partial^2 P}{\partial x^2}} + {\frac{\partial^2
      P}{\partial y^2}}\right)} = 0\;,
\label{eqn:velcon}
\end{equation}
where $t$ is the vertical time coordinate of the migrated image, $x$
and $y$ are spatial (midpoint) coordinates, and $v$ is the migration
velocity. Slightly different versions of two-dimensional implicit
extrapolation with equation (\ref{eqn:velcon}) have been described by
\cite{Li.sep.48.85} and \cite[]{Fomel.sep.92.159}. 

\plot{velcon}{width=6in}{Impulse responses of the velocity
  continuation operator, computed by an implicit, unconditionally
  stable extrapolation via the helix transform. The left plot
  corresponds to continuation towards higher velocities (migration
  mode); the right plot, smaller velocities (modeling mode).}
\par
The helix approach has allowed us to modify the old code for three
dimensions. Figure \ref{fig:velcon} shows impulse responses of an
implicit helix-based three-dimensional velocity continuation.

\sideplot{qdome}{width=3in}{Qdome synthetic model, used for testing
  the 3-D velocity continuation program.}
\par
Figure \ref{fig:modmig} illustrates the velocity continuation process
on the Qdome synthetic model \cite[]{Claerbout.gem.97}, shown in Figure
\ref{fig:qdome}. Continuation backward in velocity corresponds to the
``modeling'' mode, while forward continuation corresponds to the
``migration'' mode. It is possible to balance the amplitudes of the
two processes so that the finite-difference velocity continuation
behaves as a unitary operator
\cite[]{Fomel.sep.92.159,Fomel.sep.92.267}.

\plot{modmig}{width=6in}{Modeling (left) and migration (right) with
  the Qdome synthetic model, obtained by running the 3-D velocity
  continuation backward and forward in velocity.}

\section{Depth extrapolation and the v(x) challenge}

Can the constant-velocity result help us achieve the challenging goal
of a stable implicit depth extrapolation through media with lateral
velocity variations?
\par
The first idea that comes to mind is to replace the space-invariant
helix filters with a precomputed set of spatially varying filters,
which reflect local changes in the velocity fields. This approach
would merely reproduce the conventional practice of explicit depth
extrapolators, popularized by \cite{GPR36-02-00990114} and
\cite{GEO56-11-17701777}. However, it hides the danger of losing
the property of unconditional stability, which is obviously the major
asset of implicit extrapolators.
\par
Another route, partially explored by \cite{Nichols.sep.70.31}, is
to implement the matrix inversion in the three-dimensional implicit
scheme by an iterative method. In this case, the helix inversion may
serve as a powerful preconditioner, providing an immediate answer in
constant velocity layers and speeding up the convergence in the case
of velocity variations. To see why this might be true, one can write
the variable-coefficient matrix $\tilde{\mathbf{A}}$ in the form
\begin{equation}
  \label{eqn:golub}
  \tilde{\mathbf{A}} = \mathbf{B} + \mathbf{D}\;,
\end{equation}
where matrix $\mathbf{B}$ corresponds to some constant average velocity,
and $\mathbf{D}$ is the matrix of velocity perturbations. The system of
linear equations that we need to solve is then
\begin{equation}
  \label{eqn:system}
  \left(\mathbf{B} + \mathbf{D}\right) \mathbf{m} = \mathbf{d}\;,
\end{equation}
where $\mathbf{m}$ is the vector of extrapolated wavefield, and
$\mathbf{d}$ is an appropriate righthand side. The helix transform
provides us with the operator $\mathbf{B}^{-1}$, which we can use to
precondition system (\ref{eqn:system}). Introducing the change of
variables
\begin{equation}
  \label{eqn:prec}
  \mathbf{m} = \mathbf{B}^{-1} \mathbf{x}\;,
\end{equation}
we can transform the original system (\ref{eqn:system}) to the form
\begin{equation}
  \label{eqn:system2}
  \mathbf{d} = \left(\mathbf{B} + \mathbf{D}\right) \mathbf{B}^{-1} \mathbf{x} = 
  \mathbf{x} + \mathbf{D}\,\mathbf{B}^{-1} \mathbf{x}\;.
\end{equation}
When the velocity perturbation is small\footnote{In the
  linear-algebraic sense, this means that the spectral radius of
  operator $\mathbf{D}\,\mathbf{B}^{-1}$ is strictly less than one.}, even
the simple iteration
\begin{eqnarray}
  \label{ralg}
  \mathbf{x}_0 & = & \mathbf{d}\;; \\
  \mathbf{x}_k & = & \mathbf{d} - \mathbf{D}\,\mathbf{B}^{-1} \mathbf{x}_{k+1}\;
\end{eqnarray}
will converge rapidly to the desired solution. This interesting
possibility needs thorough testing.
\par
The third untested possibility (Papanicolaou, personal communication)
is to implement a clever patching in the velocity domain, applying a
constant-velocity filter locally inside each patch. Recently developed
fast wavelet transform techniques \cite[]{wavelet}, in particular the
\emph{local cosine transform}, provide a formal framework for that
approach.

\section{Conclusions}

The feasibility of multidimensional deconvolution, proven by the
helix transform, allows us to revisit the problem of implicit
wavefield extrapolation in three dimensions. The attraction of
implicit finite-difference methods lies in their unconditional
stability, a property invaluable for practical applications.
\par
We have shown that at least in the constant coefficient case (that is,
laterally invariant velocity), it is possible to implement an
extremely efficient implicit extrapolation by a recursive inverse
filtering in the helix-transformed computational model. Unfortunately,
the case of lateral velocity variations still presents a difficult
problem that may not have an exact solution. We are currently
exploring different roads to that goal.

\section{Acknowledgments}

We thank Biondo Biondi for useful discussions and for sharing his
three-dimensional expertise.

\bibliographystyle{seg}
\bibliography{SEP2,SEG,paper}

\append{The 1/6-th trick}
\inputdir{Math}
Given the filter $D_2 (k)$, defined in formula (\ref{eqn:d2k}), we can
construct an accurate approximation to the second derivative operator
$-k^2$ by considering a filter ratio (another Pad\'{e}-type
approximation) of the form
\begin{equation}
  \label{eqn:ratio}
  -k^2 \approx \frac{D_2(k)}{1 + \beta D_2 (k)}\;,
\end{equation}
where $\beta$ is an adjustable constant \cite[]{Claerbout.blackwell.85}.
The actual Pad\'{e} coefficient is $\beta=1/12$.  As pointed out by
Francis Muir, the value of $\beta = 1/4 - 1/\pi^2 \approx 1/6.726$
gives an exact fit at the Nyquist frequency $k = \pi$. Fitting the
derivative operator in the $L_1$ norm yields the value of $\beta
\approx 1/8.13$. All these approximations are shown in Figure
\ref{fig:sixth}.

\sideplot{sixth}{width=2.5in}{The second-derivative operator in the
  wavenumber domain and its approximations.}

\append{Constructing an ``isotropic'' Laplacian operator}
\inputdir{Sage}
The problem of approximating the Laplacian operator in two dimensions
not only inherits the inaccuracies of the one-dimensional
finite-difference approximations, but also raises the issue of
azimuthal asymmetry. For example, the usual five-point filter
\begin{equation}
\label{eqn:usual}
F_5 =
\begin{array}{|r|r|r|}
\hline
0 & 1 & 0 \\ \hline
1 & -4 & 1 \\ \hline
0 & 1 & 0  \\ \hline 
\end{array}
\end{equation}
exhibits a clear difference between the grid directions and the
directions at a 45-degree angle to the grid. To overcome this
unpleasant anisotropy, we can consider a slightly larger filter of the
form
\begin{equation}
\label{eqn:new}
F_9 = \begin{array}{|r|r|r|}
\hline
\alpha & \gamma & \alpha \\ \hline
\gamma & -4 (\alpha+\gamma) & \gamma \\ \hline
\alpha & \gamma & \alpha  \\ \hline 
\end{array}
\end{equation}
where the constants $\alpha$ and $\gamma$ are to be defined. The
Fourier-domain representation of filter (\ref{eqn:new}) is
\begin{equation}
  \label{eqn:l9k}
  F_9 (k_x,k_y) = 4\,\alpha\,[\cos{k_x}\,\cos{k_y} - 1] +
  2\,\gamma\,[\cos{k_x}+\cos{k_y}-2]\;,
\end{equation}
and the isotropic filter that we can try to approximate is defined
analogously to its one-dimensional equivalent, as follows:
\begin{equation}
  \label{eqn:lk}
  F (k_x,k_y) = 2 (\cos{k} -1) = 2 (\cos{\sqrt{k_x^2+k_y^2}} - 1)\;.
\end{equation}
Comparing equations (\ref{eqn:l9k}) and (\ref{eqn:lk}), we notice that
they match exactly, when either of the wavenumbers $k_x$ or $k_y$ is
equal to zero, provided that 
\begin{equation}
  \label{eqn:alpha}
  \alpha = \frac{1-\gamma}{2}\;.
\end{equation}
Therefore, we can reduce the problem to estimating a single
coefficient $\gamma$. Another way of expressing this conclusion is to
represent filter $F_9$ in equation (\ref{eqn:l9k}) as a linear
combination of filter $F_5$ from equation (\ref{eqn:l9k}) and its
rotated version \cite[]{cole}, as follows:
\begin{equation}
\label{eqn:new2}
F_9 = \gamma\, 
\begin{array}{|r|r|r|}
\hline
0 & 1 & 0 \\ \hline
1 & -4  & 1 \\ \hline
0 & 1 & 0  \\ \hline 
\end{array}
+ (1-\gamma)\,
\begin{array}{|r|r|r|}
\hline
1/2 & 0 & 1/2 \\ \hline
0 & -2  & 0 \\ \hline
1/2 & 0 & 1/2  \\ \hline 
\end{array}
\end{equation}
With the value of $\gamma=0.5$, filter $F_9$ takes the value
\begin{equation}
\label{eqn:mc}
F_9 = \begin{array}{|r|r|r|}
\hline
1/4 & 1/2 & 1/4 \\ \hline
1/2 & -3 & 1/2 \\ \hline
1/4 & 1/2 & 1/4  \\ \hline 
\end{array}
\end{equation}
and corresponds precisely to the nine-point McClellan filter
\cite[]{mc,GEO56-11-17781785}. On the other hand, the value of
$\gamma=2/3$ gives the least error in the vicinity of the zero
wavenumber $k$. In this case, the filter is
\begin{equation}
\label{eqn:my}
F_9 = \begin{array}{|r|r|r|}
\hline
1/6 & 2/3 & 1/6 \\ \hline
2/3 & -10/3 & 2/3 \\ \hline
1/6 & 2/3 & 1/6  \\ \hline 
\end{array}
\end{equation}
Errors of different approximations are plotted in Figure
\ref{fig:laplace}\footnote{Another way of constructing
  circular-symmetric filters is suggested by the rotated McClellan
  transform \cite[]{Biondi.sep.77.27}.}

\plot{laplace}{width=6in}{The numerical anisotropy error of different
  Laplacian approximations. Both the five-point Laplacian (plot a) and
  its rotated version (plot b) are accurate along the axes, but
  exhibit significant anisotropy in between at large wavenumbers. The
  nine-point McClellan filter (plot c) has a reduced error, while the
  filter with $\gamma=2/3$ (plot d) has the flattest error around the
  origin.}
\par

\inputdir{laplace}
Under the helix transform, a filter of the general form
(\ref{eqn:new}) becomes equivalent to a one-dimensional filter with
the $Z$ transform
\begin{eqnarray}
\label{eqn:f9z}
F_9 (Z) & = & \alpha\,Z^{-N_x-1} + \gamma\,Z^{-N_x} + \alpha\,Z^{-N_x+1} + \gamma\,Z^{-1} 
-4\,(\alpha + \gamma) \nonumber \\ & & + 
\gamma\,Z + \alpha\,Z^{N_x-1} + \gamma\,Z^{N_x} + \alpha\,Z^{N_x+1}\;,
\end{eqnarray}
where $N_x$ is the helix period (the number of grid points in the $x$
dimension). To find the inverse of a convolution with filter
(\ref{eqn:f9z}), we factorize the filter into the causal minimum-phase
component and its adjoint:
\begin{equation}
\label{eqn:factor}
F_9 (Z) = P (Z) P (1/Z)\;.
\end{equation}
To find the coefficients of the filter $P$, any one-dimensional
spectral factorization method can be applied. It is important to point
out that the result of factorization (neglecting the numerical errors)
does not depend on $N_x$. Another approach is to define a residual
error vector for the coefficients of Z in equation (\ref{eqn:factor})
and minimize it for some particular norm. For example, minimizing the
$\mathcal{L}_1$ norm when $F_9$ is the McClellan filter
(\ref{eqn:mc}), we discover that the filter $P$, after transforming
back to two dimensions, takes the form
\begin{equation}
\label{eqn:l1long}
\begin{array}{|r|r|r|r|r|r|r|r|}
\hline
     &         &           &        &     -1.6094 & 0.4293 & 0.05157 & 0.017406 \\ \hline
0.01428 & 0.033513 & 0.0808 & 0.2543 & 0.3521  & 0.1553 &   &          \\ \hline
\end{array}
\end{equation}
The results of applying a recursive deconvolution with filter
(\ref{eqn:l1long}) are shown in Figure \ref{fig:inv-laplace}.  An
essentially similar procedure, only with a different set of filters,
works for implicit wavefield extrapolation.
  
\plot{inv-laplace}{width=4in,height=4in}{Inverting the Laplacian
  operator by a helix deconvolution.  The top left plot shows the
  input, which contains a single spike and the causal minimum-phase
  filter $P$. The top right plot is the result of inverse filtering.
  As expected, the filter is deconvolved into a spike, and the spike
  turns into a smooth one-sided impulse. After the second run, in the
  backward (adjoint) direction, we obtain a numerical solution of
  Laplace's equation! In the two bottom plots, the solution is shown
  with grayscale and contours.}

%\activeplot{name}{width=6in,height=}{}{caption}
%\activesideplot{name}{height=1.5in,width=}{}{caption}
%
%\begin{equation}
%\label{eqn:}
%\end{equation}
%
%\ref{fig:}
%(\ref{eqn:})


