\author{Thomas Bayes}
%%%%%%%%%%%%%%%%%%%%%
\title{Homework 5}

\begin{abstract}
  This homework has four parts. 
  \begin{enumerate}
    \item A theoretical question related to covariance estimation.
    \item Attenuation of surface-wave noise in 
        seismic data collected over sand dunes using match filtering.
    \item Stochastic simulation of natural patterns.
    \item Missing data interpolation for ocean floor topography.
  \end{enumerate}
\end{abstract}

\section{Prerequisites}

The homework code is available from the \texttt{Madagascar} repository
by running
\begin{verbatim}
svn co https://github.com/ahay/src/trunk/book/geo384h/hw5
\end{verbatim}

\section{Theory}

Suppose that we use the gradient operator for data interpolation:
\begin{equation}
\label{eq:grad}
\min\,\left|\nabla \mathbf{m}\right|^2\;.
\end{equation}  
This approach roughly corresponds to minimizing the surface area and
represents the behavior of a soap film or a thin rubber sheet.

The corresponding inverse model covariance operator is the negative Laplacian
$\mathbf{C}_m^{-1}=\nabla^T\,\nabla=-\nabla^2$. The corresponding
covariance operator corresponds to the Green's function $G(\mathbf{x})$ that solves
\begin{equation}
\label{eq:green}
-\nabla^2 G = \delta(\mathbf{x}-\mathbf{x}_0)\;.
\end{equation}
In 2-D, the Green's function has the form 
\begin{equation}
\label{eq:g2d}
G(\mathbf{x}) = \displaystyle A - \frac{\ln |\mathbf{x}-\mathbf{x}_0|}{2\pi}
\end{equation}
with some constant $A$.

To derive equation~(\ref{eq:g2d}), we can introduce polar coordinates
around $\mathbf{x}_0$ with the radius $r= |\mathbf{x}-\mathbf{x}_0|$
and note that the Laplacian operator for a radially-symmetric function
$\phi(r)$ in polar coordinates takes the form
\begin{equation}
\label{eq:polar}
\nabla^2 \phi = \displaystyle \frac{1}{r}\,\frac{d}{dr}\,\left(r\,\frac{d \phi}{dr}\right)
\end{equation}  
Away from the point $\mathbf{x}_0$, solving
\begin{equation}
\label{eq:away}
\frac{1}{r}\,\frac{d}{dr}\,\left(r\,\frac{d G}{dr}\right) = 0
\end{equation}
leads to $G(r) = A + B\,\ln r$. To find the constant $B$, we can
integrate $\nabla^2 G$ over a circle with some small radius $\epsilon$
around the origin and apply the Green's theorem
\begin{equation}
\label{eq:integrate}
-1 = \iint \nabla^2 G dx\,dy = \oint \nabla G \cdot \vec{ds} = 
\int\limits_{0}^{2\pi} \left.\frac{\partial G}{\partial r}\right|_{r=\epsilon}\,\epsilon\,d\theta = 2\pi\,B\;.
\end{equation}

Derive the model covariance function $G(\mathbf{x})$ which corresponds
to replacing equation~(\ref{eq:grad}) with equation
\begin{equation}
\label{eq:lap}
\min\,\left|\nabla^2 \mathbf{m}\right|^2
\end{equation} 
and approximates the behavior of a thin elastic plate.


\section{Match Filtering for Attenuation of Surface Seismic Waves}
\inputdir{match}

\sideplot{data}{width=\textwidth}{Seismic shot record from sand dunes in the Middle East. The data are contaminated by ground roll propagating in the sand.}

Figure~\ref{fig:data} shows a section out of a seismic shot record
collected over sand dunes in the Middle East. The data are
contaminated by ground roll propagating in the sand. A major data
analysis task is to separate the signal (reflection waves) from the
noise (surface waves).

\plot{spec0}{width=0.8\textwidth}{Data spectrum. Solid line -- original data. Dashed line -- initial noise model and signal model.}

A quick look at the data spectrum (Figure~\ref{fig:spec0}) shows that the
noise is mostly concentrated at low frequencies. We can use this fact
to create a noise model by low-pass filtering.

\plot{noise0}{width=\textwidth}{(a) Noise model created by low-pass filtering of the original data. (b) Result of subtraction of the noise model from the data.}

Figure~\ref{fig:noise0} shows the noise model from low-pass filtering
and inner muting and the result of subtracting this model from the
data. Our next task is to match the model to the true noise by solving
the least-squares optimization problem
\begin{equation}
\label{eq:ls}
\min \|\mathbf{N}\,\mathbf{f} - \mathbf{d}\|^2\;,
\end{equation}
where $\mathbf{d}$ is the data, $\mathbf{f}$ is a \emph{matching filter}, and
$\mathbf{N}$ represents convolution of the noise model
$\mathbf{n}_0$ with the filter. After minimization, $\mathbf{n} =
\mathbf{N}\,\mathbf{f}$ becomes the new noise model, and
$\mathbf{d}-\mathbf{n}$ becomes the estimated signal. Match filtering
is implemented in program \texttt{match.c}. Some parts of this program
are left out for you to fill.

\lstset{language=c,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single,title=match/match.c]{match/match.c}

Your task:
\begin{enumerate}
\item Change directory to \texttt{hw6/match}
\item Run 
\begin{verbatim}
scons view
\end{verbatim}
to reproduce the figures on your screen.
\item Modify the \texttt{match.c} file to fill in missing parts.
\item Test your modifications by running the dot product test.
\begin{verbatim}
scons dot.test
\end{verbatim}
Repeating this several times, make sure that the numbers in the test match.
\item  Modify the \texttt{SConstruct} file to display the results of match filtering
and include them in your assignment.
\item \textbf{EXTRA CREDIT} for improving the results by finding either better parameters or a better algorithm. 
\end{enumerate}

%\newpage

\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single,title=match/SConstruct]{match/SConstruct}


\section{Natural patterns}
\inputdir{pattern}

In this section we will extract multidimensional spatial patterns from
natural images using the method of \cite{textures}. Four examples,
shown in Figures~\ref{fig:horizon}-\ref{fig:your}, contain:
\begin{enumerate}
\item Seismic horizon slice.
\item A slice from a CT-scan of a rock sample.
\item A remote-sensing satellite image.
\item Your own data (to be replaced by you).
\end{enumerate}
In each of the cases, we follow the same workflow:
\begin{enumerate}
\item Remove a linear trend from the data.
\item Estimate a multi-dimensional prediction-error filter (PEF) on a helix.
\item Apply the inverse of the estimated PEF to random normally-distributed numbers
  to create a random spatial texture, which shares the covariance with the input. 
\end{enumerate}

\plot{horizon}{width=0.8\textwidth}{Pattern extraction from a seismic time horizon.}
\plot{square}{width=0.8\textwidth}{Pattern extraction from a CT-scan of a rock sample.}
\plot{sat}{width=0.8\textwidth}{Pattern extraction from a satellite image.}
\plot{your}{width=0.8\textwidth}{Pattern extraction from your own data.}

\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single]{pattern/SConstruct}

Your task:
\begin{enumerate}
\item Change directory to \verb#geo391/hw6/pattern#
\item Run 
\begin{verbatim}
scons view
\end{verbatim}
to reproduce the figures on your screen.
\item Modify the \texttt{SConstruct} file to replace Figure~\ref{fig:your} with the figure containing your own data.
\item Why does the method fail in extracting some of the patterns? Did it succeed in extracting patterns from your data?
\end{enumerate}

\section{Missing ocean-floor data interpolation}
\inputdir{seabeam}

SeaBeam is an apparatus for measuring water depth both directly under
a boat and somewhat off to the sides of the boat's track. In this part
of the assignment, we will use a benchmark dataset from \cite{gee}:
SeaBeam data from a single day of acquisition. The original data are
shown in Figure~\ref{fig:seabeam0}a.

\plot{seabeam0}{width=0.8\textwidth}{(a) Water depth measurements from one day
  of SeaBeam acquisition. (b) Mask for locations of known data.}

Program~\texttt{interpolate.c} implements two alternative methods:
regularized inversion using convolution with a multi-dimensional
filter and preconditioning, which uses the inverse operation
(recursive deconvolution or polynomial division on a helix) for model
reparameterization.

Our first attempt is to use the Laplacian-factor filter from the
previous homework. The results from the two methods are shown in
Figure~\ref{fig:lseabeam}. They are not very successful in hiding
the ``acquisition footprint''.

\plot{lseabeam}{width=0.8\textwidth}{Left: missing data interpolation using regularization by convolution with a Laplacian factor. 
Right: missing data interpolation using model reparameterization by
deconvolution (polynomial division) with a Laplacian factor.}

Next, we will try to estimate the model covariance from the available
data. The covariance can be estimated using two alternative methods:
the fast Fourier transform (FFT) method \cite[]{jim} and the
prediction-error filter (PEF) method \cite[]{textures}. Random
realizations of model patterns from the two methods and the
corresponding 2-D Fourier spectra are shown in Figures~\ref{fig:rand2}
and~\ref{fig:rand}.

\plot{rand2}{width=0.8\textwidth}{Left: data pattern generated using FFT method for covariance estimation.
Right: its Fourier spectrum.}

\plot{rand}{width=0.8\textwidth}{Left: data pattern generated using PEF method for inverse covariance estimation.
Right: its Fourier spectrum.}


Our new attempt to interpolate missing data using the helical
prediction-error filter is shown in Figure~\ref{fig:seabeam}.

\plot{seabeam}{width=0.8\textwidth}{Left: missing data interpolation using regularization by 
convolution with a prediction-error filter. 
Right: missing data interpolation using model reparameterization by
deconvolution (polynomial division) with a prediction-error filter.}

\lstset{language=c,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single,title=seabeam/interpolate.c]{seabeam/interpolate.c}

%\newpage

\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[frame=single,title=seabeam/SConstruct]{seabeam/SConstruct}

Your task:
\begin{enumerate}
\item Change directory to \texttt{hw5/seabeam}
\item Run 
\begin{verbatim}
scons view
\end{verbatim}
to reproduce the figures on your screen.
\item Modify the \texttt{SConstruct} file to implement the following tasks:
\begin{enumerate}
\item Find the number of iterations required for both methods shown in Figure~\ref{fig:seabeam} to achieve similar results.
\item To provide a more quantitative comparison, modify the
\texttt{interpolate.c} program to output a measure of convergence
(such as the least-squares misfit) as a function of the number
of iterations\footnote{You can study the interfaces to
the \texttt{sf\_solver} and
\texttt{sf\_solver\_prec} programs. See
\url{https://sourceforge.net/p/rsf/code/HEAD/tree/trunk/api/c/bigsolver.c}}. Generate figures comparing convergence with and without preconditioning.
\end{enumerate}
\item A method for generating multiple realizations of missing data interpolation is:
\begin{enumerate}
\item Start with a random realization $\mathbf{m}_0$ such as the one shown in Figures~\ref{fig:rand2} or \ref{fig:rand}.
\item Instead of estimating $\mathbf{m}$ such that $\mathbf{K}\,\mathbf{m} = \mathbf{d}$, 
      estimate $\mathbf{x}$ such that 
\[
\mathbf{K}\,\mathbf{x} = \mathbf{d}- \mathbf{K}\,\mathbf{m_0}\;.
\]
\item The estimate for $\mathbf{m}$ is then  $\widehat{\mathbf{m}} = \widehat{\mathbf{x}}+\mathbf{m_0}$.
\end{enumerate}
Implement several realizations of missing data interpolation using
several realizations of $\mathbf{m}_0$. You can do it by modifying
either \texttt{SConstruct} or \texttt{interpolate.c}.
\item Include your results in the paper.
\item \textbf{EXTRA CREDIT} for implementing missing data interpolation using the FFT method for model covariance.
\end{enumerate}


\section{Completing the assignment}

\begin{enumerate}
\item Change directory to \texttt{hw5}.
\item Edit the file \texttt{paper.tex} in your favorite editor and change the
  first line to have your name instead of Bayes's.
\item Run
\begin{verbatim}
sftour scons lock
\end{verbatim}
to update all figures.
\item Run
\begin{verbatim}
sftour scons -c
\end{verbatim}
to remove intermediate files.
\item Run
\begin{verbatim}
scons pdf
\end{verbatim}
to create the final document.
\item Submit your result (file \texttt{paper.pdf}) on paper or by
e-mail.
\end{enumerate}

\bibliographystyle{seg}
\bibliography{hw5}
