\chapter{Writing reproducible papers, books, and reports}

Chapter~\ref{chapter:scons} describes some of the SCons tools used by
Madagascar users to write papers, books, and reports with reproducible
examples. In this chapter, we will take a closer look at an example paper.

\section{Overview of \LaTeX and SEG\TeX}

\LaTeX is a document markup language developed originally by Leslie
Lamport in early 1980s \cite[]{latex} as an extension of Donuld
Knuth's \TeX system \cite[]{tex}. \LaTeX is widely used in different
fields of science and engineering.

\LaTeX files are text files that can be edited in regular text
editors. Similarly to HTML and Markdown, \LaTeX is a descriptive
language: one can learn it by example following examples from other \LaTeX
documents. As of 2021, the Madagascar package contains nearly 300
papers and book chapters written in \LaTeX. You can find the source of
this chapter in
\href{https://github.com/ahay/src/blob/master/book/m8r/latex/paper.tex}
     {\$RSFSRC/book/m8r/latex/paper.tex}. It is also attached in the appendix.

     SEG\TeX is a collection of \LaTeX macros developed for
     geophysical publications, such as \emph{Geophysics} articles, SEG
     Annual Meeting abstracts, etc. SEG\TeX is an open-source project
     maintained by a group of volunteers. It integrates well with
     Madagascar to support the practice of \emph{reproducible
       research}, as formulated by Jon Claerbout: attaching software
     and data to the results of computational experiments to allow the
     reader to verify all computations reported in the paper using
     their own computer \cite[]{fomel2008guest}. Computational
     reproducibility is not the goal in itself but the means by which
     previous research can be extended further. It effectively turns
     all reported computational results into benchmarks and challenges
     future authors to try performing analagous computations with more
     efficiency or accuracy \cite[]{fomel2014reproducible}.

\LaTeX processing scripts included with Madagascar can turn a single
document into multiple outputs: a paper in a journal, a chapter in a
book, an expanded abstract for a conference, etc. It can also produce
multiple output formats, most importantly PDF and HTML.

\section{SCons processing}

To enable handling of \LaTeX documents, Madagascar provides special
SCons rules defined in \texttt{rsf.tex} package (for handling
individual papers) and \texttt{rsf.book} package (for handling books
and reports).

If a \LaTeX paper file has the name
\texttt{article.tex}, the \texttt{SConstruct} script for processing it
may look like
\lstset{language=python,showstringspaces=false,frame=single}
\begin{lstlisting}
  from rsf.tex import *

  Paper('article',use='amsmath',options='10pt')
\end{lstlisting}
Running
\begin{verbatim}
scons article.pdf
\end{verbatim}
processes the paper through \LaTeX to generate a PDF document.

According to the Madagascar convention, the source \LaTeX file
(\texttt{article.tex}) contains neither the usual preamble
\texttt{\\documentclass}, nor
\texttt{\\begin{document}}--\texttt{\\end{document}} commands. This is
done so that the same paper could be easily integrated into a
collection of papers in a report or as a chapter in a book.

While running the \texttt{scons article.pdf} command, SCons generates
an intermediate file \texttt{article.ltx}, which contains the
following preamble:
\lstset{language=tex,showstringspaces=false,frame=single}
\begin{lstlisting}
  \documentclass[10pt]{geophysics}
  \usepackage{amsmath}
  \begin{document}
\end{lstlisting}
  and ends with
\begin{lstlisting}
  \end{document}
\end{lstlisting}
  
Several optional arguments can be supplied to the \texttt{Paper()} command. Some of them are:
\begin{description}
\item[lclass=] specifies the LaTeX class. For example, \texttt{lclass=segabs} can be used for SEG abstracts.
\item[use=] specifies additional packages to include with \texttt{\\usepackage}. For example, \texttt{use=hyperref} can be used to add the \texttt{hyperref} package for hypertext links.
\item[include=] specifies additional commands to include in the preamble of the \LaTeX file.
\item[options=] specifies additional options for the \LaTeX class.
\end{description}
Some additional arguments control the handling of reproducible
figures and are described in the next section.

The \texttt{SConstruct} file for paper processing may end with
\begin{verbatim}
End()
\end{verbatim}
which is equivalent to setting \texttt{paper.tex} as the default
\LaTeX file. The name can be changed by giving \texttt{paper=} command
to \texttt{End()}. All other agruments to \texttt{End()} are
effectively interpreted as \verb#Paper('paper')#. The \texttt{End()}
command is not strictly necessary, but it is convenient, because
it enables command-line commands like \texttt{scons pdf},
\texttt{scons read}, etc. to be applied to the default file without
the need to spell out the name of the paper file every time.

\subsection{Including reproducible results}

A typical directory structure for writing reproducible papers follows the structure of \texttt{\$RSFSRC/book} directory in the Madagascar source repository:
\begin{verbatim}
book1/
    SConstruct [from rsf.book import *]
    paper1/
        SConstruct [from rsf.tex import *]
        paper.tex
        project1/
            SConstruct [from rsf.proj import *]
        project2/
            SConstruct [from rsf.proj import *]
\end{verbatim}

To include the results from the project directories into the paper,
first run \texttt{scons lock} in each project directory. This command
generates result figures (specified by \texttt{Result()} in the
project \texttt{SConstruct}) files and then copies them to a separate
location. The figures and all intermediate files can be then safely
removed from the current directories and data directories by running
\texttt{scons -c}. The safe directory for locked figures can be
specified by \texttt{RSFFIGS} environmental variable. By default, it is
set to \texttt{\$RSFROOT/share/madagascar/figs}.

To generate and lock all figures at once, you can run \texttt{sftour
  scons lock} from the paper directory. Analogously, all intermediate
files in project directories can be cleaned by running \texttt{sftour
  scons -c}. The \texttt{sftour} scripts simply visits all
subdirectories (whose names start with a lower-case letter) and
executes the specified command in each subdirectory.
  
Note that, when the paper is generated by \LaTeX in the paper
directory, the figures are picked from the locked directory and not
from subdirectories. This arrangement may seem counter-intituitive but
serves a clear purpose. First, it explicitly separates two activities:
running computational experiments to generate figures and writing a
paper or a book describing the results. At the end of the first
activity, the figures are locked not to be touched again. Second,
storing figures separately allows for testing and maintaining
reproducibility over time. Papers included in Madagascar store their
reproducible figures in a separate
repository\footnote{\url{https://github.com/ahay/figs}}, which is used
for a systematic testing of reproducibility, implemented in Madagascar
through continuous integration \cite[]{duvall2007continuous}.

\subsection{Multiple outputs}

Madagascar's \LaTeX conventions allow the same source to be used for
multiple different outputs: expanded abstracts, journal papers, report
papers, thesis chapters, etc. It can also produce the output in
different formats.

\subsubsection{Multiple styles}

To control the document class and the document style, use options of the \texttt{Paper} or \texttt{End} command in the paper-level \texttt{SConstruct} file, such as  \texttt{lclass=} and \texttt{use=}.

For example,
\lstset{language=python,showstringspaces=false,frame=single}
\begin{lstlisting}
  Paper('article',lclass='segabs')
\end{lstlisting}
produces an SEG expanded abstract, while
\lstset{language=python,showstringspaces=false,frame=single}
\begin{lstlisting}
  Paper('article',lclass='geophysics',options='manuscript',include='hyperref')
\end{lstlisting}
produces a Geophysics paper in the manuscript style using \textt{hyperref} package for hypertext links.

\LaTeX\ can also be used effectively for generating presentation slides. For example, the following will create slides using the \texttt{beamer} package and configuring it to use the \texttt{Madrid} theme:
\lstset{language=python,showstringspaces=false,frame=single}
\begin{lstlisting}
Paper('school10',lclass='beamer',
      include=r'''
      \usetheme{Madrid}
      ''')
\end{lstlisting}    

\subsubsection{Multiple formats}

The most common output format is PDF (Portable Document
Format). Running \texttt{scons pdf} generates a PDF file from the
\LaTeX source.

Running \texttt{scons read} generates a PDF file (if it does not exist
already) and displays it on the screen. Madagascar will search for a
PDF displaying application trying different commands in the following
order:
\begin{enumerate}
\item kpdf
\item evince
\item xpdf
\item gv
\item open 
\end{enumerate}
If none of these commands exist on your system, you need to install
one of them to be able to view PDF files.

Running \texttt{scons print} generates a PDF file (if it does not exist
already) and sends it for printing to the default printer.

Running \texttt{scons wiki} translates the \LaTeX file to the
Mediawiki format \cite[]{barrett2008mediawiki}, making it suitable for posting on the Madagascar
wiki page or other Wikipedia-style sites. The translation is crude,
done line by line other than taking the full semantic context into
account, but it can sufficient for the first draft in most cases.

Running \texttt{scons html} generates the output in HTML format using
\LaTeX2HTML, an old but powerful translator
program\footnote{\url{https://www.latex2html.org/}}. This is how
papers and book chapters included with Madagascar are published on the
website\footnote{\url{https://www.reproducibility.org/wiki/Reproducible_Documents}}. Enabling
Madagascar customizations to \LaTeX2HTML requires using the SEG\TeX
package and setting the LATEX2HTML environmental variable to
\texttt{\$TEXMF/latex2html}, where \texttt{\$TEXMF} is the location of
SEG\TeX installation (typically, \texttt{\$HOME/texmf}.) The
translation results not in a single file but in a directory of
files. By default, the name of the directory is \texttt{paper_html}.

Finally, running \texttt{scons install} packages the HTML version of
the paper together with its PDF version and all the figures for
installation on the website. By default, the packaged paper is
installed under
\texttt{\$RSFROOT/share/madagascar/book/bookname/papername}. As part
of the installation, \texttt{SConstruct} files from project
directories are converted to the HTML format to provide links for
reproducibility.

The directory structure in the installation directory make look like
the following:
\begin{verbatim}
book1/
    index.html
    ...
    paper1/
        paper.pdf
        paper_html/
            index.html
            ...
        project1.html
        project1/
            Fig/
              figure1.pdf
              figure1.png
              ...
\end{verbatim}

\subsection{Preparing papers for submission}

Madagascar attempts to simplify all steps of preparing a paper for a
publication, from numerical experiments to writing to submission. This
section describes some of the additional convenient commands defined
in \texttt{rsf.tex}.

\subsubsection{Dummy paper}

Sometimes, the figures are created before writing the paper. A
convenient command \texttt{scons dummy.tex} would generate a dummy \LaTeX
paper, which includes all the figures from subdirectories.

\subsubsection{Separate figures}

Journals often require submitting figures in separate files. To make
this step more convenient, running \texttt{scons figs} generates
figures in EPS (encapsulated postscript) format numbered according to
their appearance in the paper.

\subsubsection{Color figures}

By default, when Vplot figures are converted to EPS and PDF formats,
they are made black-and-white even if the original had colors. To
preserve colors in selected figures, use \texttt{color=} option in
\texttt{Paper} or \texttt{End} command to list the names of figures
that should have their colors preserved. For example, the following
\texttt{SConstruct} file
\lstset{language=python,showstringspaces=false,frame=single}
\begin{lstlisting}
  from rsf.tex import *

  Paper('article',use='amsmath',options='10pt',
           color='map velocity')
\end{lstlisting}
will preserve color in figures named \texttt{map} and
\texttt{velocity}. You can also use \texttt{color='ALL'} to keep the
color in all figures.

\subsubsection{High-resolution figures}

To story bitmap images (such as those produced by Madagascar's
\texttt{sfgrey} command), PDF uses  a lossy compression algorithm,
which can blur some of the finer details. To switch to lossless
compression (which generates larger files but keeps the initial
resolution intact), use \texttt{hires=} option in
\texttt{Paper} or \texttt{End} command. For example, the following
\texttt{SConstruct} file
\lstset{language=python,showstringspaces=false,frame=single}
\begin{lstlisting}
  from rsf.tex import *

  Paper('article',use='amsmath',options='10pt',
           color='map velocity',
           hires='velocity image')
\end{lstlisting}
will preserve color in figures named \texttt{map} and
\texttt{velocity} and will preserve high resolution in figures named
\texttt{velocity} and \texttt{image}.

\subsubsection{Non-reproducible figures}

\subsubsection{Non-Madagascar figures}

\section{Including papers into books and reports}

The production of books and reports using Madagascar tools is
controlled by customized SCons commands specified in the
\texttt{rsf.book} package.

An example \texttt{SConstruct} file at the book level may look like
the following:
\lstset{language=python,showstringspaces=false,frame=single}
\begin{lstlisting}
  from rsf.book import *

  Book(title=''My Book'',author=''My Name'',
          chapters=''intro theory discussion'')

  End(options='book')
\end{lstlisting}
The \texttt{chapters=} argument in the \texttt{Book} command contains
the list of chapter directories in the order of their appearance in
the book.

\subsection{Different kinds of books}

\subsubsection{Book}

\subsubsection{Report}

\subsubsection{Thesis}

\section{Multiple outputs}

% pdf

% www

\section{Acknowledgments}

The \LaTeX-processing scripts in Madagascar were directly inspired by
a system developed earlier at SEP (Stanford Exploration Project) for
implementing Jon Claerbout's vision of reproducible research
\cite[]{sep}.

\bibliographystyle{seg}
\bibliography{latex}

\appendix
\section{Appendix: The source of this document}
\verbatiminput{paper.tex}

