\chapter{Vplot graphics}

The final results of a data analysis project are often represented by figures. Madagascar includes a versatile plotting library for generating scientific illustrations called Vplot. Vplot defines a data format for storing figures. Figures stored in that format can be displayed on the screen or transformed into other formats for inclusion in scientific papers, presentation slides, or other documents.

\section{Vplot history}

Vplot was developed in 1980s at the Stanford Exploration Project
(SEP). At that time, there was no universal format for graphics files
and no univeral libraries for representing graphics on hardware
devices, such as monitors or printers. The purpose of the Vplot was to
provide such a format so that the same plots could be easily
transferred between devices
\cite[]{Cole.sep.60.349,Dellinger.sep.61.327}.


The \texttt{man} page for Vplot says
  \begin{quote}
    Vplot originally stood for `vector plot', but this name is now
    insufficient because vplot supports not only vectors, but also
    filled areas (either raster patterns or hatched), text (Hershey
    fonts), and raster images (including grey-scale dithered ones on
    monochrome devices such as laser printers).
    \end{quote}

Joe Dellinger was the main developer of Vplot, other major
contributors at Stanford were Steve Cole and Dave Nichols, with
additional contributions by other SEP students. After Vplot was
adopted by Madagascar, new filter programs (called ``pens'') for
converting Vplot files to other formats were added.

\section{Vplot format}
\inputdir{test}

Vplot breaks all graphics into primitives that are encoded in the file
in a mixed text and binary form. The source of the Vplot library is
contained in the Madagascar source tree under
\texttt{\$RSFSRC/plot/lib} with example main functions in
\texttt{\$RSFSRC/plot/test}.

For example, here is a portion of a function that draws a diamond
shape to reproduce a figure (Figure~\ref{fig:croshyp}) from \cite[]{bei}:

\lstset{language=c,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[firstline=106,lastline=128,frame=single,title=\$RSFSRC/plot/test/croshyp.c]{\RSF/plot/test/croshyp.c}

The function calls are self-descriptive: \texttt{vp\_umove} sets the
starting point (in predefined ``user'' coordinates),
\texttt{vp\_udraw} draws a line to the next point. Four lines form a
diamond shape.

\sideplot{croshyp}{width=\textwidth}{A figure reproduced from (Claerbout, 2006)
  shows an example of using Vplot library.}

A documentation of the programming interface to different functions in
the library can be obtained by running \texttt{man libvplot}.

\subsection{Python interface}

It is also possible to generate Vplot files from Python using the
Python interface to the library defined in \texttt{vplot} module.

For example, here is an equivalent (slighly more concise) Python code
for generating a diamond shape in Figure ~\ref{fig:croshyp}:

\lstset{language=python,numbers=left,numberstyle=\tiny,showstringspaces=false}
\lstinputlisting[firstline=32,lastline=42,frame=single,title=\$RSFSRC/plot/test/croshyp.py]{\RSF/plot/test/croshyp.py}

\subsection{sfpldb and sfplas}

Vplot files use a mixture of binary and text code, which is not easy
to decipher, Two convenient programs, \texttt{sfpldb} and
\texttt{sfplas}, allow the user to convert a Vplot file to a pure text representation and back to Vplot.

For a quick demonstration, let us create an example Vplot file:
\begin{verbatim}
bash$ sfspike n1=10 k1=5 | sfsmooth rect1=3 | sfgraph title=Example > test.vpl
\end{verbatim}

Suppose that we want to change the axis label from ``Time (s)'' to ``Distance (km)'' without regenerating the file. This can be done by
\begin{enumerate}
\item converting the Vplot file to a text file
\begin{verbatim}
bash$ < test.vpl sfpldb > test.txt
\end{verbatim}
\item editing the text file \texttt{test.txt}
\item converting back to Vplot
\begin{verbatim}
bash$ < test.txt sfplas > test2.vpl
\end{verbatim}
\end{enumerate}

Aternatively, we can do it using a Unix pipe and \texttt{sed} text editor, as follows:
\begin{verbatim}
bash$ < test.vpl sfpldb | sed s/Time\ \(s\)/Distance\ \(km\)/ | sfplas > test2.vpl
\end{verbatim}

\subsection{sfvplotdiff and reproducibility testing}

One of the key reasons for a continued use of Vplot in Madagascar is
the need for reproducibility testing. To test reproducibility, one can
repeat the computation, generate a new figure and compare it with the
previously stored figure. The \texttt{sfvplotdiff} provides a smart
comaprison, which either confirms that the two figures match (up to
a certain threshold) or identifies the mismatch.

Let us try a simple test by creating severak slightly different figures.
\begin{verbatim}
bash$ sfspike n1=5 k1=3 | sfgraph title=Hello > fig1.vpl
bash$ sfspike n1=5 k1=3 | sfgraph title=Bonjour > fig2.vpl
bash$ sfspike n1=5 k1=3 | sfnoise var=1e-8 | sfgraph title=Hello > fig3.vpl
bash$ sfspike n1=5 k1=3 | sfnoise var=1e-6 | sfgraph title=Hello > fig4.vpl
\end{verbatim}

Comparison by \texttt{sfvplotdiff} correctly identifies that the
difference between \texttt{fig1.vpl} and \texttt{fig2.vpl} is in the
title text.
\begin{verbatim}
bash$ sfvplotdiff fig1.vpl fig2.vpl 
sfvplotdiff: Text string mismatch
	File fig1.vpl, line 91: Hello
	File fig2.vpl, line 91: Bonjour
\end{verbatim}

Comparing \texttt{fig1.vpl} and \texttt{fig3.vpl},
\texttt{sfvplotdiff} doesn't report any difference and returns without
messages. The data changed, but the difference was insignificant and
not noticeable in the plot.
\begin{verbatim}
bash$ sfvplotdiff fig1.vpl fig3.vpl
\end{verbatim}
On the other hand, comparing \texttt{fig1.vpl} and \texttt{fig4.vpl},
\texttt{sfvplotdiff} reports noticeable differences in the plot:
\begin{verbatim}
sfvplotdiff fig1.vpl fig4.vpl
sfvplotdiff: Coordinate mismatch: draw
	File fig1.vpl, lines 69 and 70: draw from (1365,3515) to (1293,3515)
	File fig4.vpl, lines 69 and 70: draw from (1365,3518) to (1293,3518)
sfvplotdiff: State mismatch: move
	File fig1.vpl, line 72 (G): move=(1238,3515) from line 71.
	File fig4.vpl, line 72 (G): move=(1238,3518) from line 71.
sfvplotdiff: Coordinate mismatch: draw
	File fig1.vpl, lines 74 and 75: draw from (1365,4401) to (1293,4401)
	File fig4.vpl, lines 74 and 75: draw from (1365,4404) to (1293,4404)
[... more output ...]
\end{verbatim}

Comparisons by \texttt{sfvplotdiff} are incorporated in the automatic
testings scripts used by Madagascar to test reproducibility and to
maintain the functionality of the package through continuous integration.

\section{Graphics programs}

In this section, we will look at examples of some of the most popular
graphics programs. These programs take a data file (in the RSF format)
as input and produce a figure (in the Vplot format) as output.

\subsection{sfdots}

\subsection{sfgraph}
c
\subsection{sfgraph3}

\subsection{sfwiggle}

\subsection{sfbargraph}

\subsection{sfcontour}

\subsection{sfcontour3}

\subsection{sfthplot}

\subsection{sfgrey}

\subsection{sfgrey3}

\subsection{sfgrey4}

\subsection{sfbox}

\section{Pen programs}

\subsection{xtpen}

\subsection{oglpen}

\subsection{pspen}

\section{Vplot and interactivity}

\bibliographystyle{seg}
\bibliography{SEP2,vplot}
